---
title: "Heritable transgenerational fitness variation correlates with copper resistance in the clonal duckweed Spirodela polyrhiza"
author: "Alexandra Chávez"
date: 05.05.2025 
doi: "https://doi.org/10.1098/rspb.2025.3304"
---
################################################
##### Clean workspace and upload libraries #####
################################################

```{r}
rm(list=ls()) 
set.seed(123)
```

```{r}
 ## for stats
library(glmmTMB)
library(DHARMa)

## for plots
library(ggplot2) ## for plots

##for data
library(readxl) #upload excel file
library(dplyr) #states the grammar of data manipulation
library(data.table) ## inserting data :=
library(psych) ## for descriptive stats
library(plyr) ## for replacing values in a variable
library(ggpmisc)# for R2 values in ggplot
library(performance)# for R2 values for partial models

##other packages
library(inti)## an alternative for calculating heritability
#library(knitr)# create tables from the inti output
library(sommer)#for h2 
#remotes::install_github("mastoffel/partR2") 
library(partR2) ##heritability analysis

##for volcano plots
library (ggrepel)## to exclude labels in ggplot

theme_set(theme_classic(base_size = 18))
```


##############################################################
##############################################################
##############################################################
###### Fitness and phenotype anaylsis of 56 genotype #########
##############################################################
##############################################################
##############################################################


```{r}
df<- read_excel("./GeneticBase_TransgnerationalPlasticity.xlsx", sheet = "56genotypes",col_names = T, na = "")
```
  # convert to factors
```{r}
df$Pretreat<-factor(df$Pretreat)
df$TreatEnv<-factor(df$TreatEnv)
df$Population<-factor(df$Population)
df$Genotype<-factor(df$Genotype)
df$Sample<-factor(df$Sample)
df$Replicate<-factor(df$Replicate)

df$Pretreat1<-factor(df$Pretreat, levels=c("Cu-","Cu+"), ordered=TRUE)
df$TreatEnv1<-factor(df$TreatEnv, levels=c("Control","Copper"), ordered=TRUE)
df$Population1<-factor(df$Population, levels = c("America","Europe","India","Asia"), ordered=T)

gen=unique(df$Genotype)
env=unique(df$TreatEnv)
pre=unique(df$Pretreat)
rep=unique(df$Replicate)

df2<-df
```

########

#############################################################
##### Variance Analysis  of Population Growth Rate ##########
#############################################################

## 1 - Variance in whole data
```{r}
dfq2.1<-df2[complete.cases(df2$GRArea),]

dfq2.1%>%dplyr::group_by(TreatEnv,Pretreat)%>%dplyr::summarise(N=length(GRArea))

car::leveneTest(data=dfq2.1[dfq2.1$TreatEnv=="Control",],GRArea~(Pretreat)) ##F test #0.06638
car::leveneTest(data=dfq2.1[dfq2.1$TreatEnv=="Copper",],GRArea~(Pretreat)) ##F test #0.0004447

dfq2.1%>%dplyr::group_by(TreatEnv,Pretreat)%>%dplyr::summarise(data_mean=mean(GRArea, na.rm=T),N=length(GRArea))

pdf("./Fig1b.pdf",height=6,width=4)
ggplot(dfq2.1,aes(Pretreat, GRArea,fill=Pretreat))+
  geom_violin(aes(col=Pretreat),alpha=0.5,width=1)+facet_wrap(.~TreatEnv)+
  geom_boxplot(width=0.5)+guides(col="none",fill="none")+
  scale_color_manual(values = c("Cu-"="#0056B3", "Cu+"="#CC0000"))+
  scale_fill_manual(values = c("Cu-"="#0056B3", "Cu+"="#CC0000"))+
  scale_y_continuous(breaks=seq(0,0.6,by=0.2),limits = c(0,0.6), expand = c(0,0))+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
        axis.text.y = element_text(size = 25,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        legend.text = element_text(size = 20,color="black", family = "sans"),
        legend.background = element_rect(fill = F),
        legend.key = element_rect(fill = F),
        legend.title = element_text(size = 30,color="black", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```

## 2 - perform Kolmogorov-Smirnov test
```{r}
# Differences of skewness
df2.a<-df2[complete.cases(df2$GRArea),]
ks.test(df2.a[df2.a$TreatEnv=="Control"&df2.a$Pretreat=="Cu-",]$GRArea,
        df2.a[df2.a$TreatEnv=="Control"&df2.a$Pretreat=="Cu+",]$GRArea) 

ks.test(df2.a[df2.a$TreatEnv=="Copper"&df2.a$Pretreat=="Cu-",]$GRArea,
        df2.a[df2.a$TreatEnv=="Copper"&df2.a$Pretreat=="Cu+",]$GRArea) 

# Skewness
statpsych::test.skew(df2.a[df2.a$TreatEnv=="Control"&df2.a$Pretreat=="Cu-",]$GRArea)
statpsych::test.skew(df2.a[df2.a$TreatEnv=="Control"&df2.a$Pretreat=="Cu+",]$GRArea)
statpsych::test.skew(df2.a[df2.a$TreatEnv=="Copper"&df2.a$Pretreat=="Cu-",]$GRArea)
statpsych::test.skew(df2.a[df2.a$TreatEnv=="Copper"&df2.a$Pretreat=="Cu+",]$GRArea)

df2.a[df2.a$TreatEnv=="Control"&df2.a$GRArea<0.36,]

pdf("./FigS2d.pdf",height=6,width=8)
ggplot(df2.a,aes(GRArea))+
  geom_density(alpha=0.5, aes(fill=Pretreat))+
  facet_grid(~TreatEnv)+  
  guides(fill="none")+
  scale_x_continuous(breaks=seq(0,0.6,by=0.2),limits=c(0,0.6))+
  scale_y_continuous(limits=c(0,15),expand=c(0,0))+
  scale_fill_manual (values = c("Cu-"="#0056B3","Cu+"="#CC0000"))+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
        axis.text.y = element_text(size = 18,color="black", family = "sans"),
        axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
        legend.text = element_text(size = 20,color="black", family = "sans"),
        legend.background = element_rect(fill = F),
        legend.key = element_rect(fill = F),
        legend.title = element_text(size = 25,color="black", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```

###  3 - estimate variance manually for analysis = df2Var 
```{r}
df2Var<-(df2 %>% dplyr::group_by(Pretreat,TreatEnv,Genotype,Population) %>% dplyr::summarise(VGRArea= var(GRArea, na.rm = TRUE)))

df2Var<-data.table(as.data.frame(df2Var[complete.cases(df2Var$Genotype),]))

df2Var%>%dplyr::group_by(TreatEnv,Pretreat)%>%dplyr::summarise(data_mean=mean(VGRArea, na.rm=T),N=length(VGRArea))

## General plot
a<-round(kruskal.test(data=df2Var[TreatEnv=="Control",], VGRArea~Pretreat)$p.value,3)
b<-round(kruskal.test(data=df2Var[TreatEnv=="Copper",],  VGRArea~Pretreat)$p.value,4)
ann_text<-data.frame(label=paste("Kruskal(P) =",c(a,b)),TreatEnv=c("Control","Copper"))

pdf("./Fig1c.pdf",height=6,width=4)
ggplot(df2Var, aes(x=Pretreat, y=VGRArea))+
  geom_violin(aes(fill=Pretreat),alpha=0.5)+
  geom_boxplot(width=0.1,aes(fill=Pretreat))+
  labs(x="Environment Treatment", y="Variance per Genotype")+
  scale_y_continuous(breaks=seq(0,0.02,by=0.005),limits=c(0,0.02), expand = c(0,0))+
  scale_color_manual(values = c("Cu-"="#0056B3","Cu+"="#CC0000"))+
  scale_fill_manual (values = c("Cu-"="#0056B3","Cu+"="#CC0000"))+
  guides(fill="none",color="none")+
  facet_grid(.~TreatEnv)+
  geom_text(data=ann_text,mapping=aes(x=1.5,y=0.015,label=label))+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
        axis.text.y = element_text(size = 25,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        legend.text = element_text(size = 20,color="black", family = "sans"),
        legend.background = element_rect(fill = F),
        legend.key = element_rect(fill = F),
        legend.title = element_text(size = 30,color="black", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```

### 4 - Growth rate smaller than 70%
```{r}
df1_1<-data.table(df[,c(1:4,7)])
df1_1<-df1_1[complete.cases(df1_1),]

list1<-data.frame(Genotype=character(),Replicate=character(),Pretreat=character(),TreatEnv=character(),Area=numeric(),Reference=numeric(),stringsAsFactors=FALSE)
for (q in gen){
    for (e in pre){
      table<-df1_1[Genotype==q&Pretreat==e&TreatEnv=="Copper",]
      a<-as.numeric(with(table, median(GRArea, na.rm = TRUE)))
      for(i in 1:nrow(table)){
        if (table[i,]$GRArea<=(a*0.7)){
          list1<-rbind(list1,data.frame(Genotype=q,
                                        Replicate=table[i,]$Replicate,
                                        Pretreat=e,
                                        TreatEnv=table[i,]$TreatEnv,
                                        Area=table[i,]$GRArea,
                                        Reference=a))  }}
      }}
list1
list1$Genotype ##cu-: "15"  "26"  "32"  "72"  "110" "133" "143" "172" "181" "194" "197" "223" "227", cu+: "110" "142" "179" 

#Plot
pdf("./Fig1d.pdf",height=6,width=4)
table2<-data.table(count=c(0,0,13,3),TreatEnv=rep(c("Control","Copper"),each=2),Pretreat=rep(c("Cu-","Cu+"),2))
ggplot(table2,aes(Pretreat,count,fill=Pretreat))+
  geom_col()+
  facet_grid(~TreatEnv)+
  scale_fill_manual (values = c("Cu-"="#0056B3","Cu+"="#CC0000"))+
  guides(fill="none",color="none")+
  scale_y_continuous(breaks=seq(0,14,by=2),limits=c(0,14), expand = c(0,0))
dev.off()
```
 #proportion test for smaller growth
```{r}
#GRArea
prop.test(x=c(10,0),  n=c(56,56)) #60
prop.test(x=c(13,3),  n=c(56,56)) 
prop.test(x=c(13,3), n=c(56,56)) #70
prop.test(x=c(16,3), n=c(56,56))
prop.test(x=c(20,10),n=c(56,56))#80
```

### 5 - Variance among genotypes
```{r}
df2Average<-(df2 %>% dplyr::group_by(Genotype,TreatEnv,Pretreat,Population) %>% dplyr::summarise(MeanGRArea   = mean(GRArea, na.rm = TRUE)))

df2Average<-data.table(df2Average)
df2Average<-df2Average[complete.cases(df2Average),]
df2Average%>%dplyr::group_by(TreatEnv,Pretreat)%>%dplyr::summarise(N=length(MeanGRArea))

## Variance analysis of the mean
car::leveneTest(data=df2Average[df2Average$TreatEnv=="Control",],MeanGRArea~(Pretreat)) ##F test
car::leveneTest(data=df2Average[df2Average$TreatEnv=="Copper",], MeanGRArea~(Pretreat)) ##F test

pdf("./FigS2b.pdf",height=6,width=4)
ggplot(df2Average, aes(y=MeanGRArea, x=Pretreat))+
  geom_violin(aes(col=Pretreat,fill=Pretreat),alpha=0.5,width=1)+
  geom_boxplot(width=0.5,aes(fill=Pretreat))+
  facet_wrap(.~TreatEnv)+
  guides(col="none",fill="none")+
  scale_color_manual(values = c("Cu-"="#0056B3", "Cu+"="#CC0000"))+
  scale_fill_manual(values = c("Cu-"="#0056B3", "Cu+"="#CC0000"))+
  labs(y="MeanGRArea/Genotype")+
  scale_y_continuous(breaks=seq(0,0.5,by=0.1),limits = c(0.2,0.5), expand = c(0,0))+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
        axis.text.y = element_text(size = 25,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        legend.text = element_text(size = 20,color="black", family = "sans"),
        legend.background = element_rect(fill = F),
        legend.key = element_rect(fill = F),
        legend.title = element_text(size = 30,color="black", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```

## 6 - perform Kolmogorov-Smirnov test for mean values
```{r}
# Differences of skewness
df2.a<-df2Average[complete.cases(df2Average$MeanGRArea),]
ks.test(df2.a[df2.a$TreatEnv=="Control"&df2.a$Pretreat=="Cu-",]$MeanGRArea,
        df2.a[df2.a$TreatEnv=="Control"&df2.a$Pretreat=="Cu+",]$MeanGRArea) 

ks.test(df2.a[df2.a$TreatEnv=="Copper"&df2.a$Pretreat=="Cu-",]$MeanGRArea,
        df2.a[df2.a$TreatEnv=="Copper"&df2.a$Pretreat=="Cu+",]$MeanGRArea) 

# Skewness
statpsych::test.skew(df2.a[df2.a$TreatEnv=="Control"&df2.a$Pretreat=="Cu-",]$MeanGRArea)
statpsych::test.skew(df2.a[df2.a$TreatEnv=="Control"&df2.a$Pretreat=="Cu+",]$MeanGRArea)
statpsych::test.skew(df2.a[df2.a$TreatEnv=="Copper"&df2.a$Pretreat=="Cu-",]$MeanGRArea)
statpsych::test.skew(df2.a[df2.a$TreatEnv=="Copper"&df2.a$Pretreat=="Cu+",]$MeanGRArea)

df2.a[df2.a$TreatEnv=="Control"&df2.a$MeanGRArea<0.36,]

pdf("./FigS2a.pdf",height=6,width=8)
ggplot(df2.a,aes(MeanGRArea))+
  geom_density(alpha=0.5, aes(fill=Pretreat))+
  facet_grid(~TreatEnv)+  
  guides(fill="none")+
  scale_x_continuous(breaks=seq(0,0.6,by=0.1),limits=c(0.2,0.5))+
  scale_y_continuous(limits=c(0,25),expand=c(0,0))+
  scale_fill_manual (values = c("Cu-"="#0056B3","Cu+"="#CC0000"))+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
        axis.text.y = element_text(size = 18,color="black", family = "sans"),
        axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
        legend.text = element_text(size = 20,color="black", family = "sans"),
        legend.background = element_rect(fill = F),
        legend.key = element_rect(fill = F),
        legend.title = element_text(size = 25,color="black", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```


###############################
#### Pre-treatment Effect #####
###############################

3) Do not run this section if using the pre-treatment ratios of 16 genotypes
  If you already did, rm(list=ls()) and reload the data

### creating values of effect of copper pre-treatment
```{r}
## Pre-treatment effect
df3<-data.table(df2)
for (x in gen){
for (y in env){
        a<-as.numeric(with(df3[Genotype==x&TreatEnv==y&Pretreat=="Cu-",],mean(GRArea, na.rm = TRUE)))
	    	                   df3[Genotype==x&TreatEnv==y&Pretreat=="Cu+", paste('PreGRArea') := GRArea/a]
        b<-as.numeric(with(df3[Genotype==x&TreatEnv==y&Pretreat=="Cu-",],mean(OffFvFm, na.rm = TRUE)))
	    	                   df3[Genotype==x&TreatEnv==y&Pretreat=="Cu+", paste('PreOffFvFm') := OffFvFm/b]
}}
```

## Creating pre-treatment clean data set 
```{r}
df4<- df3[Pretreat=="Cu+",]

## Create new replicate variable
df4 <- do.call(data.frame,lapply(df4,function(x) replace(x, is.infinite(x), NA)))
df4<-data.table(df4)
```

### Mean of each genotype for the main plots = df4Mean
```{r}
df4Mean<-(df4 %>% dplyr::group_by(Genotype,TreatEnv,Population) %>% dplyr::summarise(
  MPreGRArea    = mean(PreGRArea, na.rm = TRUE), 
  MPreOffFvFm = mean(PreOffFvFm, na.rm = TRUE),
  ))
df4Mean<-data.table(df4Mean)
```

## Mean data values
```{r}
dfMean<-(df2 %>% dplyr::group_by(Genotype,Pretreat,TreatEnv,Population) %>% dplyr::summarise(
  MGRArea    = mean(GRArea, na.rm = TRUE)   , 
  MOffFvFm = mean(OffFvFm, na.rm = TRUE), 
  ))
```

### Data of pre-treatment ratios copper vs control
```{r}
dfq2.2.1<-merge((df4Mean[df4Mean$TreatEnv=="Copper",]),(df4Mean[df4Mean$TreatEnv=="Control",]),by=c("Genotype","Population")) #"Country",
dfq2.2.1<- dfq2.2.1[,-c(3,6)]
colnames(dfq2.2.1)<-c("Genotype","Population", 
                      "PreGRArea.cu","PreFvFm.cu",
                      "PreGRArea.ctr","PreFvFm.ctr")
```

###############################
#### Resistance Ratios ########
###############################

1) Do not run this section if using resistance from the 16 genotypes 
      If you already did, rm(list=ls()) and reload the data
      
### creating values of effect of copper induction (first time stress)
```{r}
## Pre-treatment effect
dfR0<-data.table(df2)
for (x in gen){
for (y in pre){
        a<-as.numeric(with(dfR0[Genotype==x&Pretreat==y&TreatEnv=="Control",], mean(GRArea, na.rm = TRUE)))
	    	                   dfR0[Genotype==x&Pretreat==y&TreatEnv=="Copper", paste('ReGRArea') := GRArea/a]
}}
```

## Creating resistance clean data set 
```{r}
dfR<- dfR0[TreatEnv=="Copper",]

## Create new replicate variable
dfR <- do.call(data.frame,lapply(dfR,function(x) replace(x, is.infinite(x), NA)))
dfR<-data.table(dfR)
```

### Mean of each genotype for the analysis
```{r}
dfRMean<-(dfR %>% dplyr::group_by(Genotype,Pretreat,Population) %>% dplyr::summarise(MReGRArea    = mean(ReGRArea, na.rm = TRUE)))
dfRMean<-data.table(dfRMean)
```

2) Do not run this section if using growth rate of naive plants of 16 genotypes
      If you already did, rm(list=ls()) and reload the data
      
### Creating GRArea of naive plants
```{r}
dfNaive<-df2[df2$Pretreat=="Cu-",]
dfNaive<-dfNaive%>%dplyr::group_by(Genotype,TreatEnv)%>%dplyr::summarise(NaiveGRArea=mean(GRArea, na.rm=T))
dfNaiveCtr<-dfNaive[dfNaive$TreatEnv=="Control",]
dfNaiveCu<-dfNaive[dfNaive$TreatEnv=="Copper",]
```

3) Do not run this section if using the pre-treatment ratios of 16 genotypes
      If you already did, rm(list=ls()) and reload the data

### Merge pre-treatment and resistance data
```{r}
dfRMeanCu<-dfRMean[Pretreat=="Cu-",]
df4MeanCu<-df4Mean[TreatEnv=="Copper",]
df4MeanCtr<-df4Mean[TreatEnv=="Control",]

ResPreCu<-merge(dfRMeanCu,df4MeanCu,by=c("Genotype","Population"))
ResPreCu<-merge(ResPreCu,dfNaiveCu,by=c("Genotype","TreatEnv"))

ResPreCtr<-merge(dfRMeanCu,df4MeanCtr,by=c("Genotype","Population"))
ResPreCtr<-merge(ResPreCtr,dfNaiveCtr,by=c("Genotype","TreatEnv"))
```

### Merge data under control and copper conditions and resistance or pre-treatment effect
```{r}
### For control plants against resistance
dfMeanCtr<-dfMean[dfMean$TreatEnv=="Control"&dfMean$Pretreat=="Cu-",]
dfRMeanCu<-dfRMean[Pretreat=="Cu-",]

CtrRe<-merge(dfMeanCtr,dfRMeanCu,by=c("Genotype","Population","Pretreat"))

### For control plants against pre-treatment
dfMeanCtr<-dfMean[dfMean$TreatEnv=="Control"&dfMean$Pretreat=="Cu-",]
df4MeanCtr<-df4Mean[TreatEnv=="Control",]
df4MeanCu<-df4Mean[TreatEnv=="Copper",]

CtrPreCtr<-merge(dfMeanCtr,df4MeanCtr,by=c("Genotype","Population"))
CtrPreCu<-merge(dfMeanCtr,df4MeanCu,by=c("Genotype","Population"))

#### errase this section
### For plants under copper against pre-treatment
dfMeanCu<-dfMean[dfMean$TreatEnv=="Copper"&dfMean$Pretreat=="Cu-",]
df4MeanCtr<-df4Mean[TreatEnv=="Control",]
df4MeanCu<-df4Mean[TreatEnv=="Copper",]

CuPreCtr<-merge(dfMeanCu,df4MeanCtr,by=c("Genotype","Population"))
CuPreCu<-merge(dfMeanCu,df4MeanCu,by=c("Genotype","Population"))
```





##################################################
##################################################
##################################################
###### Repetition of 16 genotypes ################
##################################################
##################################################
##################################################




#upload data of metabolites
```{r}
df_Rep<-read_excel("./GeneticBase_TransgnerationalPlasticity.xlsx", col_names = T, na = "",sheet="Repetition")
```

# convert to factors
```{r}
df_Rep$TreatEnv<-factor(df_Rep$TreatEnv)
df_Rep$Pretreat<-factor(df_Rep$Pretreat)
df_Rep$Pretreat1<-factor(df_Rep$Pretreat, levels=c("Cu-","Cu+"), ordered=TRUE)
df_Rep$TreatEnv1<-factor(df_Rep$TreatEnv, levels=c("Control","Copper"), ordered=TRUE)
df_Rep$Genotype<-factor(df_Rep$Genotype)
df_Rep$Population<-factor(df_Rep$Population)
df_Rep$DaughterPretreat<-factor(df_Rep$DaughterPretreat)
df_Rep$DaughterTreatEnv<-factor(df_Rep$DaughterTreatEnv)

env=unique(df_Rep$TreatEnv)
pre=unique(df_Rep$Pretreat)
rep=unique(df_Rep$Replicate)
gen=unique(df_Rep$Genotype)
```

# Assess the effect of using the first and second daughter
```{r}
options(contrasts = c("contr.treatment", "contr.poly"))
summary(lm(data=df_Rep, GRArea~DaughterPretreat))
summary(lm(data=df_Rep, GRArea~DaughterTreatEnv))
```


###############################
### Pre-treatment Effect 
###############################

```{r}
df_Rep2<-data.table(df_Rep); dim(df_Rep2)

for (x in gen){
  for (y in env){
   a<-as.numeric(with(df_Rep2[Genotype==x&TreatEnv==y&Pretreat=="Cu-",], mean(GRArea, na.rm = TRUE)))
	    	              df_Rep2[Genotype==x&TreatEnv==y&Pretreat=="Cu+", paste('PreMeanGRArea') := GRArea/a]
}}
df_Rep3<-df_Rep2[Pretreat=="Cu+",]
```

 #function to extract beta estimates
```{r}
## https://www.researchgate.net/post/Can-we-compare-betas-of-two-different-regression-analyses
    #glmmTMB model
stdCoef.merModTMB <- function(object,data) {
  sdy <- sd(data)# the y values are now in the 'y' slot of the resp attribute
  sdx <- apply(getME(lm2,"X"), 2, sd) # And the X matriz is in the 'X' slot of the pp attr
  sc <- fixef(object)[[1]]*sdx/sdy #beta coefficient = usual regression coefficient multiplied by the sample standard deviation of IV(X) and divide by the sample standard deviation of the DV (Y). 
  se.fixef <- coef(summary(object))$cond[,"Std. Error"] # last change - extracting the standard errors from the summary
  se <- se.fixef*sdx/sdy
  return(data.frame(stdcoef=sc, stdse=se))
}
```
      #creating data for PreGRArea  ~Genotype*TreatEnv
```{r}
#Data
df_Rep4<-df_Rep3[complete.cases(df_Rep3$PreMeanGRArea),]

## Keep only genotypes that were repeated
df4<-df4[complete.cases(df4$PreGRArea),]
df5<-df4[df4$Genotype %in% gen,]

#Models
options(contrasts = c("contr.treatment", "contr.poly"))
lm2<-glmmTMB(data=df_Rep4, PreMeanGRArea~Genotype*TreatEnv+(1|Genotype:Replicate)+(1|Rack),family=Gamma(log))
lm2<-glmmTMB(data=df_Rep4, PreMeanGRArea~Genotype*TreatEnv+(1|Genotype:Replicate)+(1|Rack),dispformula = ~TreatEnv)
sim1<-simulateResiduals(lm2,n=500, plot=T);testDispersion(sim1);testZeroInflation(sim1)
plotResiduals(lm2,df_Rep4$Genotype);plotResiduals(lm2,df_Rep4$TreatEnv)

lm4<-glmmTMB(data=df5, PreGRArea~Genotype*TreatEnv+(1|Genotype:Sample)+(1|Rack))
sim1<-simulateResiduals(lm4,n=500, plot=T);testDispersion(sim1);testZeroInflation(sim1)
plotResiduals(lm4,df5$Genotype);plotResiduals(lm4,df5$TreatEnv)

#transformation using the functions
dat2<-stdCoef.merModTMB(lm2,df_Rep4$PreMeanGRArea)
dat4<-stdCoef.merModTMB(lm4,df5$PreGRArea)

data1whole<-cbind(dat2,dat4);
colnames(data1whole)<-c("stdcoef.lm2","stdse.lm2","stdcoef.lm4","stdse.lm4")
data1whole$TreatEnv<-c(rep("Control",16),rep("Copper",16))

options(contrasts = c("contr.treatment", "contr.poly")) ## default
model<-glmmTMB(data=data1whole,stdcoef.lm4~stdcoef.lm2*TreatEnv)
simulateResiduals(model,n=500, plot=T)
summary(model)
car::Anova(model,type=2)
options(contrasts = c("contr.sum", "contr.poly")) ## for type III
modelp<-glmmTMB(data=data1whole,stdcoef.lm4~stdcoef.lm2*TreatEnv)
car::Anova(modelp,type=3)

pdf("./Fig5.pdf",height=6,width=6)
ggplot(data1whole,aes(y=stdcoef.lm4,x=stdcoef.lm2,color=TreatEnv))+
  geom_line(aes(y=predict(model, type="response")),size=1.5)+
  geom_point(size=8)+
  guides(color="none")+
  scale_x_continuous(breaks=seq(-0.3,0.3,by=0.2),limits=c(-0.3,0.3),expand=c(0,0))+
  scale_y_continuous(breaks=seq(-0.25,0.75,by=0.25),limits=c(-0.25,0.75),expand=c(0,0))+
  scale_color_manual(values = c("Control"="#0056B3", "Copper"="#CC0000"))+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()
```



#############################################################
### correlation between the first and second experiment #####
#############################################################

1) When using resistance of 16 genotypes versus growth rate of naive plants and pre-treatment ratios of 56 genotypes
### to extract 16 genotype plant resistance to first time stress
```{r}
dfR0<-data.table(df_Rep)
for (x in gen){
for (y in pre){
        a<-as.numeric(with(dfR0[Genotype==x&Pretreat==y&TreatEnv=="Control",], mean(GRArea, na.rm = TRUE)))
	    	                   dfR0[Genotype==x&Pretreat==y&TreatEnv=="Copper", paste('ReGRArea') := GRArea/a]
}}

dfR<- dfR0[TreatEnv=="Copper",]

## Create new replicate variable
dfR <- do.call(data.frame,lapply(dfR,function(x) replace(x, is.infinite(x), NA)))
dfR<-data.table(dfR)

dfRMean<-(dfR %>% dplyr::group_by(Genotype,Pretreat,Population) %>% dplyr::summarise(MReGRArea    = mean(ReGRArea, na.rm = TRUE)))
dfRMean<-data.table(dfRMean)
```

### Merge 56 genotype pre-treatment and 16 genoytpe resistance data
```{r}
#Fitness of naive plants
dfNaive<-df2[df2$Pretreat=="Cu-",]
dfNaive<-dfNaive%>%dplyr::group_by(Genotype,TreatEnv)%>%dplyr::summarise(NaiveGRArea=mean(GRArea, na.rm=T))
dfNaiveCtr<-dfNaive[dfNaive$TreatEnv=="Control",]
dfNaiveCu<-dfNaive[dfNaive$TreatEnv=="Copper",]

dfRMeanCu<-dfRMean[Pretreat=="Cu-",]
df4MeanCu<-df4Mean[TreatEnv=="Copper",]
df4MeanCtr<-df4Mean[TreatEnv=="Control",]

ResPreCu<-merge(dfRMeanCu,df4MeanCu,by=c("Genotype","Population"))
ResPreCu<-merge(ResPreCu,dfNaiveCu,by=c("Genotype","TreatEnv"))

ResPreCtr<-merge(dfRMeanCu,df4MeanCtr,by=c("Genotype","Population"))
ResPreCtr<-merge(ResPreCtr,dfNaiveCtr,by=c("Genotype","TreatEnv"))
```

### Merge fitness of naive plants and resistance data
```{r}
### For control plants against resistance
dfMeanCtr<-dfMean[dfMean$TreatEnv=="Control"&dfMean$Pretreat=="Cu-",]
dfMeanCu<-dfMean[dfMean$TreatEnv=="Copper"&dfMean$Pretreat=="Cu-",]
dfRMeanCu<-dfRMean[Pretreat=="Cu-",]

CtrRe<-merge(dfMeanCtr,dfRMeanCu,by=c("Genotype","Population","Pretreat"))
```

2) When using the growth rate of naive plants from 16 genotypes vs the resistance and pre-treatment of 56 genotypes
### Creating GRArea of naive plants
```{r}
dfNaive<-df_Rep[df_Rep$Pretreat=="Cu-",]
dfNaive<-dfNaive%>%dplyr::group_by(Genotype,TreatEnv)%>%dplyr::summarise(MGRArea=mean(GRArea, na.rm=T))
dfNaiveCtr<-dfNaive[dfNaive$TreatEnv=="Control",]
dfNaiveCu<-dfNaive[dfNaive$TreatEnv=="Copper",]
```

### Merge data under control and copper conditions and resistance or pre-treatment effect
```{r}
### For control plants against resistance
dfRMeanCu<-dfRMean[Pretreat=="Cu-",]

CtrRe<-merge(dfNaiveCtr,dfRMeanCu,by=c("Genotype"))

### For 16 genotypes naive plants against 56 genotype pre-treatment respective to the treatment
df4MeanCtr<-df4Mean[TreatEnv=="Control",]
df4MeanCu<-df4Mean[TreatEnv=="Copper",]

CtrPreCtr<-merge(dfNaiveCtr,df4MeanCtr,by=c("Genotype","TreatEnv"))
CtrPreCu<-merge(dfNaiveCtr,df4MeanCu,by=c("Genotype"))
CtrPreCuCopper<-merge(dfNaiveCu,df4MeanCu,by=c("Genotype","TreatEnv"))
```

3) When using the pre-treatment of 16 genotypes vs growth rate of naive plants and resistance of 56 genotypes
### creating values of effect of copper pre-treatment again for the correlations
```{r}
## Pre-treatment effect
df3<-data.table(df_Rep)
for (x in gen){
for (y in env){
        a<-as.numeric(with(df3[Genotype==x&TreatEnv==y&Pretreat=="Cu-",],mean(GRArea, na.rm = TRUE)))
	    	                   df3[Genotype==x&TreatEnv==y&Pretreat=="Cu+", paste('PreGRArea') := GRArea/a]
}}

df4<- df3[Pretreat=="Cu+",]

## Create new replicate variable
df4 <- do.call(data.frame,lapply(df4,function(x) replace(x, is.infinite(x), NA)))
df4<-data.table(df4)

## Mean values per genotype
df4Mean<-(df4 %>% dplyr::group_by(Genotype,TreatEnv,Population) %>% dplyr::summarise(
  MPreGRArea    = mean(PreGRArea, na.rm = TRUE)  ))
df4Mean<-data.table(df4Mean)
```

### Merge pre-treatment and resistance data
```{r}
dfRMeanCu<-dfRMean[Pretreat=="Cu-",]
df4MeanCu<-df4Mean[TreatEnv=="Copper",]
df4MeanCtr<-df4Mean[TreatEnv=="Control",]

ResPreCu<-merge(dfRMeanCu,df4MeanCu,by=c("Genotype","Population"))

ResPreCtr<-merge(dfRMeanCu,df4MeanCtr,by=c("Genotype","Population"))
```

### Merge data under control and copper conditions and resistance or pre-treatment effect
```{r}
dfMean<-(df2 %>% dplyr::group_by(Genotype,Pretreat,TreatEnv,Population) %>% dplyr::summarise(
  MGRArea    = mean(GRArea, na.rm = TRUE) ))

dfMeanCtr<-dfMean[dfMean$TreatEnv=="Control"&dfMean$Pretreat=="Cu-",]
dfMeanCu<-dfMean[dfMean$TreatEnv=="Copper"&dfMean$Pretreat=="Cu-",]

### For 56 genotypes naive plants against 16 genotype pre-treatment respective to the treatment
df4MeanCtr<-df4Mean[TreatEnv=="Control",]
df4MeanCu<-df4Mean[TreatEnv=="Copper",]

CtrPreCtr<-merge(dfMeanCtr,df4MeanCtr,by=c("Genotype","TreatEnv","Population"))
CtrPreCu<-merge(dfMeanCtr,df4MeanCu,by=c("Genotype","Population"))
CtrPreCuCopper<-merge(dfMeanCu,df4MeanCu,by=c("Genotype","TreatEnv","Population"))

```

###############################################
######  Pre-treatment effect analysis #########
###############################################

## General
## Assess the effect of the genotype and the pre-treatment on the fitness
```{r}
dfq2.2<-df4[complete.cases(df4$PreGRArea),]

options(contrasts = c("contr.treatment", "contr.poly"))

lm1<-glmmTMB(data=dfq2.2, PreGRArea~Genotype*TreatEnv+(1|Rack),family=Gamma(log),dispformula=~TreatEnv,control=glmmTMBControl(optCtrl=list(iter.max=1e5,eval.max=1e5)))
sim1<-simulateResiduals(lm1,n=500, plot=T);testDispersion(sim1);testZeroInflation(sim1)
plotResiduals(lm1,dfq2.2$Genotype);plotResiduals(lm1,dfq2.2$TreatEnv)
summary(lm1)

options(contrasts = c("contr.sum", "contr.poly"))
modelp<-glmmTMB(data=dfq2.2, PreGRArea~Genotype*TreatEnv+(1|Rack),family=Gamma(log),dispformula=~TreatEnv,control=glmmTMBControl(optCtrl=list(iter.max=1e5,eval.max=1e5)))
car::Anova(modelp,type=3)

dfq2.2%>%dplyr::group_by(TreatEnv)%>%dplyr::summarise(data_mean=mean(PreGRArea, na.rm=T),N=length(PreGRArea))
length(df4Mean[df4Mean$TreatEnv=="Copper"&df4Mean$MPreGRArea>1,]$Genotype)
length(df4Mean[df4Mean$TreatEnv=="Copper"&df4Mean$MPreGRArea<1,]$Genotype)
(25*100)/56
(31*100)/56

df4Mean2<-(df4 %>% dplyr::group_by(TreatEnv,Genotype) %>% dplyr::summarise(
  MPreGRArea   = mean(PreGRArea,na.rm=TRUE),
  se = (sd(PreGRArea,na.rm=TRUE)/sqrt(sum(!is.na(PreGRArea))))))
df2PreGRArea <-df4Mean2[complete.cases(df4Mean2$MPreGRArea),]
order.df<-df2PreGRArea[df2PreGRArea$TreatEnv=="Copper",]
order.df<-order.df[order(order.df$MPreGRArea),]
list<-unique(order.df$Genotype)
df2PreGRArea$Genotype<-factor(df2PreGRArea$Genotype,ordered = T,levels=(list))

pdf(".s/FigS3.pdf",height=6,width=16)
ggplot(df2PreGRArea, aes(x=Genotype, y=MPreGRArea))+
  geom_point(aes(col=TreatEnv),size=3)+
  geom_errorbar(aes(col=TreatEnv,ymin=MPreGRArea-se, ymax=MPreGRArea+se),width=0.6,size=0.4)+
  labs(x="Environment Treatment")+
  geom_hline(yintercept = 1,linetype="dashed",color="black", size=0.6)+
  scale_color_manual(values = c("Control"="#0056B3","Copper"="#CC0000"))+
  guides(color="none")+
  facet_grid(~TreatEnv)+
  scale_y_continuous(breaks=seq(0.5,1.5,by=0.25),limits = c(0.5,1.5), expand = c(0,0))+
  theme(axis.text.y = element_text(size=25,vjust = 1, hjust=0.8,color="black", family = "sans"),
     axis.text.x = element_text(size = 5,color="black", family = "sans",angle = 90),
     axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
     legend.text = element_text(size = 20,color="black", family = "sans"),
     legend.background = element_rect(fill = F),
     legend.key = element_rect(fill = F),
     legend.title = element_text(size = 30,color="black", family = "sans"),
     axis.ticks = element_line(color="black")) 
dev.off()
```
       # Correlation pre-treatment ratio in copper and control (mean data)
```{r}
dfq2.2.1<-dfq2.2.1[complete.cases(dfq2.2.1$PreGRArea.cu),]
options(contrasts = c("contr.treatment", "contr.poly"))

lm01<-glmmTMB(data=dfq2.2.1, PreGRArea.ctr~PreGRArea.cu*Population)
lm01<-glmmTMB(data=dfq2.2.1, PreGRArea.ctr~PreGRArea.cu+(1|Population))
sim<-simulateResiduals(lm01,n=500, plot=T);testDispersion(sim)
summary(lm01)
car::Anova(lm01,type=2)

pred_df <- data.frame(PreGRArea.cu = seq(min(dfq2.2.1$PreGRArea.cu), max(dfq2.2.1$PreGRArea.cu), length.out = 100))
pred_df$PreGRArea.ctr <- predict(lm01, newdata = pred_df, re.form = NA)

pdf("./Fig2a.pdf",height=6,width=6)
 ggplot(dfq2.2.1, aes(y = PreGRArea.ctr, x = PreGRArea.cu)) + 
  geom_point(color="black", size=5) + 
  geom_hline(yintercept=1,linetype="dashed")+
  geom_vline(xintercept=1,linetype="dashed")+
  theme(axis.text = element_text(size=16), axis.title = element_text(size=20))+
  geom_line(data=pred_df, aes(y = PreGRArea.ctr, x = PreGRArea.cu),color="#CC0000",size=2)+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  scale_y_continuous(breaks=seq(0.9,1.2,by=0.1),limits = c(0.9,1.2),expand=c(0,0))+
  scale_x_continuous(breaks=seq(0.7,1.45,by=0.15),limits = c(0.7,1.45),expand=c(0,0))+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.ticks = element_line(color="black"))
 dev.off()

 sample<-factor(c("4","8","32","35","39","43","69","72","103","126","186","197","198","207","212","235"))
 df_labels <- dfq2.2.1[dfq2.2.1$Genotype %in% sample, ]
 
 pdf("./FigS8.pdf",height=6,width=6)
 ggplot(dfq2.2.1, aes(y = PreGRArea.ctr, x = PreGRArea.cu)) + 
  geom_point(color="black", size=5) + 
  geom_hline(yintercept=1,linetype="dashed")+
  geom_vline(xintercept=1,linetype="dashed")+
  geom_text(data = df_labels, aes(label = Genotype), vjust = -1, size = 5,family = "sans") +
  theme(axis.text = element_text(size=16), axis.title = element_text(size=20))+
  geom_line(data=pred_df, aes(y = PreGRArea.ctr, x = PreGRArea.cu),color="#CC0000",size=2)+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  scale_y_continuous(breaks=seq(0.9,1.2,by=0.1),limits = c(0.9,1.2),expand=c(0,0))+
  scale_x_continuous(breaks=seq(0.7,1.45,by=0.15),limits = c(0.7,1.45),expand=c(0,0))+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.ticks = element_line(color="black"))
 dev.off()
```
        # histogram plot for heritability
```{r}
dfq2.2<-df4[complete.cases(df4$PreGRArea),]

df4Mean2<-(df4 %>% dplyr::group_by(TreatEnv,Genotype) %>% dplyr::summarise(
  MPreGRArea   = mean(PreGRArea,na.rm=TRUE),
  se = (sd(PreGRArea,na.rm=TRUE)/sqrt(sum(!is.na(PreGRArea))))))
df2PreGRArea <-df4Mean2[complete.cases(df4Mean2$MPreGRArea),]

df2PreGRArea%>%dplyr::group_by(TreatEnv)%>%dplyr::summarise(
  data_mean=mean(MPreGRArea, na.rm=T),
  max=max(MPreGRArea, na.rm=T),
  min=min(MPreGRArea, na.rm=T),
  N=length(MPreGRArea))

pdf("./Fig2a.pdf",height=6,width=8)
ggplot(df2PreGRArea, aes(x=round(MPreGRArea,2)))+
  geom_histogram(binwidth=0.05, aes(fill=TreatEnv))+#
  facet_grid(~TreatEnv)+
  scale_fill_manual (values = c("Control"="#0056B3","Copper"="#CC0000"))+
  scale_x_continuous(breaks=seq(0.75,1.5,by=0.25),limits=c(0.72,1.5),expand=c(0,0))+
  scale_y_continuous(breaks=seq(0,25,by=5),limits=c(0,25),expand=c(0,0))+
  geom_vline(xintercept = 1,linetype="dashed",color="black", size=1)+
  guides(fill="none")+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()
```


       # correlation between resistance of first time stressed plants and pre-treatment
```{r}
# Pre-treatment in control
model<-glmmTMB(data=ResPreCtr,MPreGRArea~MReGRArea+NaiveGRArea+(1|Population))
sim<-simulateResiduals(model,n=500, plot=T)
plotResiduals(model,ResPreCtr$MPreGRArea);plotResiduals(model,ResPreCtr$NaiveGRArea)
summary(model)
car::Anova(model)

model<-glmmTMB(data=ResPreCtr,MPreGRArea~MReGRArea+NaiveGRArea+(1|Population))
newdat <- data.frame(MReGRArea = seq(min(ResPreCtr$MReGRArea), max(ResPreCtr$MReGRArea), length = 100),
                     NaiveGRArea = mean(ResPreCtr$NaiveGRArea, na.rm = TRUE), Population=mean(ResPreCtr$Population))
newdat$pred <- predict(model, newdata = newdat, type = "response")
partR2( model,  partvars = c("MReGRArea", "NaiveGRArea"), data = ResPreCtr )

pdf("./Fig3a.pdf",height=6,width=4)
ggplot(data=ResPreCtr,aes(y=MPreGRArea,x=MReGRArea))+
  geom_point(size=3)+
  labs(x="Resistance under first time stress",y="Pre-treatment ratio under control")+
  geom_line(data = newdat,aes(x = MReGRArea, y = pred),size=1.5,col="#CC0000")+
  scale_y_continuous(breaks=seq(0.9,1.2,by=0.1),limits=c(0.9,1.2),expand=c(0,0))+
  scale_x_continuous(breaks=seq(0.5,1,by=0.1),limits=c(0.5,1),expand=c(0,0))+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()

# Pre-treatment in copper
model<-glmmTMB(data=ResPreCu,MPreGRArea~MReGRArea+NaiveGRArea+(1|Population))
sim<-simulateResiduals(model,n=500, plot=T)
plotResiduals(model,ResPreCu$MPreGRArea);plotResiduals(model,ResPreCu$NaiveGRArea)
summary(model)
car::Anova(model)

model<-glmmTMB(data=ResPreCu,MPreGRArea~MReGRArea+NaiveGRArea+(1|Population))
newdat <- data.frame(MReGRArea = seq(min(ResPreCu$MReGRArea), max(ResPreCu$MReGRArea), length = 100),
                     NaiveGRArea = mean(ResPreCu$NaiveGRArea, na.rm = TRUE), Population=mean(ResPreCu$Population))
newdat$pred <- predict(model, newdata = newdat, type = "response")
partR2(model,  partvars = c("MReGRArea", "NaiveGRArea"), data = ResPreCtr )

pdf("./Fig3b.pdf",height=6,width=4)
ggplot(data=ResPreCu,aes(y=MPreGRArea,x=MReGRArea))+
  geom_point(size=3)+
  labs(x="Resistance under first time stress",y="Pre-treatment ratio under copper")+
  geom_line(data = newdat,aes(x = MReGRArea, y = pred),size=1.5,col="#CC0000")+
  scale_y_continuous(breaks=seq(0.7,1.5,by=0.2),limits=c(0.7,1.5),expand=c(0,0))+
  scale_x_continuous(breaks=seq(0.5,1,by=0.1),limits=c(0.5,1),expand=c(0,0))+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()
```
       # correlation between 16genotype resistance of first time stressed plants and 56 genotype pre-treatment
```{r}
# Pre-treatment in control
model<-glmmTMB(data=ResPreCtr,MPreGRArea~MReGRArea+(1|Population))
sim<-simulateResiduals(model,n=500, plot=T)
plotResiduals(model,ResPreCtr$MPreGRArea)
summary(model)
car::Anova(model)

newdat <- data.frame(MReGRArea = seq(min(ResPreCtr$MReGRArea), max(ResPreCtr$MReGRArea), length = 100),
                     Population=mean(ResPreCtr$Population))
newdat$pred <- predict(model, newdata = newdat, type = "response")

pdf("./FigS10a1.pdf",height=6,width=4)
ggplot(data=ResPreCtr,aes(y=MPreGRArea,x=MReGRArea))+
  geom_point(size=3)+
  labs(x="Resistance under first time stress\n16 genotypes",y="Pre-treatment ratio under control\n56 genotypes")+
  geom_line(data = newdat,aes(x = MReGRArea, y = pred),size=1.5,col="#CC0000")+
  scale_y_continuous(breaks=seq(0.9,1.2,by=0.1),limits=c(0.9,1.2),expand=c(0,0))+
  scale_x_continuous(breaks=seq(0.68,0.84,by=0.04),limits=c(0.68,0.84),expand=c(0,0))+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()

# Pre-treatment in copper
model<-glmmTMB(data=ResPreCu,MPreGRArea~MReGRArea+(1|Population))
sim<-simulateResiduals(model,n=500, plot=T)
plotResiduals(model,ResPreCu$MPreGRArea)
summary(model)
car::Anova(model)

newdat <- data.frame(MReGRArea = seq(min(ResPreCu$MReGRArea), max(ResPreCu$MReGRArea), length = 100),
                     Population=mean(ResPreCu$Population))
newdat$pred <- predict(model, newdata = newdat, type = "response")

pdf("./FigS10b1.pdf",height=6,width=4)
ggplot(data=ResPreCu,aes(y=MPreGRArea,x=MReGRArea))+
  geom_point(size=3)+
  labs(x="Resistance under first time stress\n16 genotypes",y="Pre-treatment ratio under copper\n56 genotypes")+
  geom_line(data = newdat,aes(x = MReGRArea, y = pred),size=1.5,col="#CC0000")+
  scale_y_continuous(breaks=seq(0.7,1.5,by=0.2),limits=c(0.7,1.5),expand=c(0,0))+
  scale_x_continuous(breaks=seq(0.68,0.84,by=0.04),limits=c(0.68,0.84),expand=c(0,0))+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()
```
       # correlation between 56genotype resistance of first time stressed plants and 16 genotype pre-treatment
```{r}
# Pre-treatment in control
model<-glmmTMB(data=ResPreCtr,MPreGRArea~MReGRArea+(1|Population))
sim<-simulateResiduals(model,n=500, plot=T)
plotResiduals(model,ResPreCtr$MPreGRArea)
summary(model)
car::Anova(model)

newdat <- data.frame(MReGRArea = seq(min(ResPreCtr$MReGRArea), max(ResPreCtr$MReGRArea), length = 100),
                     Population=mean(ResPreCtr$Population))
newdat$pred <- predict(model, newdata = newdat, type = "response")

pdf("./FigS10a2.pdf",height=6,width=4)
ggplot(data=ResPreCtr,aes(y=MPreGRArea,x=MReGRArea))+
  geom_point(size=3)+
  labs(x="Resistance under first time stress\n56 genotypes",y="Pre-treatment ratio under control\n16 genotypes")+
  geom_line(data = newdat,aes(x = MReGRArea, y = pred),size=1.5,col="#CC0000")+
  scale_y_continuous(breaks=seq(0.95,1.07,by=0.03),limits=c(0.95,1.07),expand=c(0,0))+
  scale_x_continuous(breaks=seq(0.5,1,by=0.15),limits=c(0.5,0.95),expand=c(0,0))+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()

# Pre-treatment in copper
model<-glmmTMB(data=ResPreCu,MPreGRArea~MReGRArea+(1|Population))
sim<-simulateResiduals(model,n=500, plot=T)
plotResiduals(model,ResPreCu$MPreGRArea)
summary(model)
car::Anova(model)

newdat <- data.frame(MReGRArea = seq(min(ResPreCu$MReGRArea), max(ResPreCu$MReGRArea), length = 100),
                     Population=mean(ResPreCu$Population))
newdat$pred <- predict(model, newdata = newdat, type = "response")

pdf("./FigS10b2.pdf",height=6,width=4)
ggplot(data=ResPreCu,aes(y=MPreGRArea,x=MReGRArea))+
  geom_point(size=3)+
  labs(x="Resistance under first time stress\n56 genotypes",y="Pre-treatment ratio under copper\n16 genotypes")+
  geom_line(data = newdat,aes(x = MReGRArea, y = pred),size=1.5,col="#CC0000")+
  scale_y_continuous(breaks=seq(0.9,1.1,by=0.05),limits=c(0.9,1.1),expand=c(0,0))+
  scale_x_continuous(breaks=seq(0.5,1,by=0.15),limits=c(0.5,0.95),expand=c(0,0))+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()
```

       # correlation between 56 genotype growth rate under control and 16 genotype plant resistance for first time
```{r}
modelplot<-glmmTMB(data=CtrRe,MReGRArea~MGRArea) ## model doesnt adjust so we considered the correlation with the inverted values
model<-glmmTMB(data=CtrRe,MGRArea~MReGRArea+(1|Population))
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model)

pdf("./FigS11b.pdf",height=6,width=4)
ggplot(data=CtrRe,aes(y=MReGRArea,x=MGRArea))+
  geom_point(size=3)+
  labs(x="Growth under control\n56 genotypes",y="Resistance under first time stress\n16 genotypes")+
  geom_line(aes(y=predict(modelplot, type="response")),size=1.5,col="#CC0000")+
  scale_x_continuous(breaks=seq(0.34,0.49,by=0.05),limits=c(0.34,0.49),expand=c(0,0))+
  scale_y_continuous(breaks=seq(0.68,0.84,by=0.04),limits=c(0.68,0.84),expand=c(0,0))+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()
```
       # correlation between 16 genotype growth rate under control and 56 genotype plant resistance for first time
```{r}
modelplot<-glmmTMB(data=CtrRe,MGRArea~MReGRArea,family=Gamma(log))
modelplot<-glmmTMB(data=CtrRe,MReGRArea~MGRArea)
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(modelplot)

pdf("./FigS11a.pdf",height=6,width=4)
ggplot(data=CtrRe,aes(y=MReGRArea,x=MGRArea))+
  geom_point(size=3)+
  labs(x="Growth under control\n16 genotypes",y="Resistance under first time stress\n56 genotypes")+
  geom_line(aes(y=predict(modelplot, type="response")),size=1.5,col="#CC0000")+
  scale_y_continuous(breaks=seq(0.5,1,by=0.15),limits=c(0.5,0.95),expand=c(0,0))+
  scale_x_continuous(breaks=seq(0.34,0.49,by=0.05),limits=c(0.34,0.49),expand=c(0,0))+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()
```

       # correlation between growth rate under control and plant pre-treatment under copper
```{r}
model<-glmmTMB(data=CtrPreCu,MPreGRArea~MGRArea+(1|Population),family=Gamma(log))
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model)

pdf("./FigS12c.pdf",height=6,width=4)
ggplot(data=CtrPreCu,aes(y=MPreGRArea,x=MGRArea))+
  geom_point(size=3)+
  labs(x="Growth under control",y="Pre-treatment ratio under copper")+
  geom_line(aes(y=predict(model, type="response")),size=1.5,col="#CC0000")+
  scale_y_continuous(breaks=seq(0.7,1.5,by=0.2),limits=c(0.7,1.5),expand=c(0,0))+
  scale_x_continuous(breaks=seq(0.34,0.49,by=0.05),limits=c(0.34,0.49),expand=c(0,0))+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()
```
       # correlation between 16 genotype growth rate of naive plants and 56 genotype plant pre-treatment
```{r}
model<-glmmTMB(data=CtrPreCtr,MPreGRArea~MGRArea+(1|Population))
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model)

pdf("./FigS12a1.pdf",height=6,width=4)
ggplot(data=CtrPreCtr,aes(y=MPreGRArea,x=MGRArea))+
  geom_point(size=3)+
  labs(x="Growth under control\n16 genotypes",y="Pre-treatment ratio under control\n56 genotypes")+
  geom_line(aes(y=predict(model, type="response")),size=1.5,col="#CC0000")+
  scale_y_continuous(breaks=seq(0.9,1.2,by=0.1),limits=c(0.9,1.2),expand=c(0,0))+
  scale_x_continuous(breaks=seq(0.34,0.49,by=0.05),limits=c(0.34,0.49),expand=c(0,0))+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()


model<-glmmTMB(data=CtrPreCu,MPreGRArea~MGRArea+(1|Population))
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model)

pdf("./FigS12b1.pdf",height=6,width=4)
ggplot(data=CtrPreCu,aes(y=MPreGRArea,x=MGRArea))+
  geom_point(size=3)+
  labs(x="Growth under control\n16 genotypes",y="Pre-treatment ratio under copper\n56 genotypes")+
  geom_line(aes(y=predict(model, type="response")),size=1.5,col="#CC0000")+
  scale_y_continuous(breaks=seq(0.7,1.5,by=0.2),limits=c(0.7,1.5),expand=c(0,0))+
  scale_x_continuous(breaks=seq(0.34,0.49,by=0.05),limits=c(0.34,0.49),expand=c(0,0))+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()

model<-glmmTMB(data=CtrPreCuCopper,MGRArea~MPreGRArea,dispformula = ~Population)
model<-glmmTMB(data=CtrPreCuCopper,MGRArea~MPreGRArea+(1|Population))
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model)

ggplot(data=CtrPreCuCopper,aes(x=MPreGRArea,y=MGRArea))+
  geom_point(size=3)+
  labs(y="Growth under copper\n16 genotypes",x="Pre-treatment ratio under copper\n56 genotypes")+
  geom_line(aes(y=predict(model, type="response")),size=1.5,col="#CC0000")+
  #scale_x_continuous(breaks=seq(0.7,1.5,by=0.2),limits=c(0.7,1.5),expand=c(0,0))+
  #scale_y_continuous(breaks=seq(0.34,0.49,by=0.05),limits=c(0.34,0.49),expand=c(0,0))+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
```
       # correlation between 56 genotype growth rate of naive plants and 16 genotype plant pre-treatment
```{r}
model<-glmmTMB(data=CtrPreCtr,MPreGRArea~MGRArea+(1|Population))
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model)

pdf("./FigS12a2.pdf",height=6,width=4)
ggplot(data=CtrPreCtr,aes(x=MGRArea,y=MPreGRArea))+
  geom_point(size=3)+
  labs(x="Growth under control\n56 genotypes",y="Pre-treatment ratio under control\n16 genotypes")+
  geom_line(aes(y=predict(model, type="response")),size=1.5,col="#CC0000")+
  scale_x_continuous(breaks=seq(0.34,0.49,by=0.05),limits=c(0.34,0.49),expand=c(0,0))+
  scale_y_continuous(breaks=seq(0.96,1.08,by=0.04),limits=c(0.96,1.08),expand=c(0,0))+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()

model<-glmmTMB(data=CtrPreCu,MPreGRArea~MGRArea)
model<-glmmTMB(data=CtrPreCu,MPreGRArea~MGRArea+(1|Population)) #doesnt adjust
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model)

pdf("./FigS12b2.pdf",height=6,width=4)
ggplot(data=CtrPreCu,aes(y=MPreGRArea,x=MGRArea))+
  geom_point(size=3)+
  labs(x="Growth under control\n56 genotypes",y="Pre-treatment ratio under copper\n16 genotypes")+
  geom_line(aes(y=predict(model, type="response")),size=1.5,col="#CC0000")+
  scale_y_continuous(breaks=seq(0.9,1.11,by=0.07),limits=c(0.9,1.11),expand=c(0,0))+
  scale_x_continuous(breaks=seq(0.34,0.49,by=0.05),limits=c(0.34,0.49),expand=c(0,0))+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()
```
   
##########################################################
###### GRArea and FvFm correlation analysis ##############
##########################################################

    ## Create clean datasets
```{r}
df4_clean<-df4[complete.cases(df4$PreGRArea),]
df4_clean<-df4_clean[complete.cases(df4_clean$PreOffFvFm),]

df4Mean_clean<-df4Mean[complete.cases(df4Mean$MPreGRArea),]
df4Mean_clean<-df4Mean_clean[complete.cases(df4Mean_clean$MPreOffFvFm),]
```
   #Modelling with mean values per genotype
```{r}
options(contrasts = c("contr.treatment", "contr.poly"))
mod1<-glmmTMB(data=df4Mean_clean,MPreGRArea~MPreOffFvFm*TreatEnv+(1|Genotype),dispformula = ~TreatEnv)
sim<-simulateResiduals(mod1,n=500, plot=T);testDispersion(sim)
summary(mod1)

options(contrasts = c("contr.sum", "contr.poly")) 
modp<-glmmTMB(data=df4Mean_clean,MPreGRArea~MPreOffFvFm*TreatEnv+(1|Genotype),dispformula = ~TreatEnv)
car::Anova(modp,test="Chi",type=3)

modCtr<-glmmTMB(data=df4Mean_clean[df4Mean_clean$TreatEnv=="Control",],MPreGRArea~MPreOffFvFm)
modCu <-glmmTMB(data=df4Mean_clean[df4Mean_clean$TreatEnv=="Copper",], MPreGRArea~MPreOffFvFm)
sim<-simulateResiduals(modCtr,n=500, plot=T);testDispersion(sim)
sim<-simulateResiduals(modCu, n=500, plot=T);testDispersion(sim)

summary(modCtr)
summary(modCu)
car::Anova(modCtr,test="Chi")
car::Anova(modCu, test="Chi")

pdf("./Fig4a.pdf",height=6,width=6)
ggplot(df4Mean_clean, aes(y=MPreGRArea, x=MPreOffFvFm, color=TreatEnv))+
  geom_point(size=2)+
  geom_smooth(method='lm',se=F,size=1.5,formula = y~x)+
  stat_poly_eq(formula = x~y, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  scale_color_manual(values=c("Control"="#0056B3", "Copper"="#CC0000"))+
  scale_x_continuous(breaks=seq(0.95,1.15,by=0.05),limits = c(0.94,1.15), expand = c(0,0))+
  scale_y_continuous(breaks=seq(0.7,1.5,by=0.2),limits = c(0.7,1.5), expand = c(0,0))+
  guides(color="none")+
  theme(axis.text.x = element_text(size=18,vjust = 1, hjust=0.8,color="black", family = "sans"),
     axis.text.y = element_text(size = 18,color="black", family = "sans"),
     axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
     legend.text = element_text(size = 20,color="black", family = "sans"),
     legend.background = element_rect(fill = F),
     legend.key = element_rect(fill = F),
     legend.title = element_text(size = 30,color="black", family = "sans"),
     axis.ticks = element_line(color="black")) 
dev.off()
```


#################################
###### Heritability  ############
#################################

  ## Analysis with Cullis_H2
      ## Create function
```{r}
Cullis_H2TMB <- function(model) {
  # extract genetic variance
  var_g <- glmmTMB::VarCorr(model, comp = "Variance")[[1]]$Genotype[1] 
  # retrieve conditional variance of BLUPs
  convars <- glmmTMB::ranef(model, condVar = TRUE)  #returns the conditional variance of Best Linear Unbiased Predictions (BLUPs) for Genotype
  g_convar <- attr(convars[[1]]$Genotype, "condVar") #captures the conditional variance for each Genotype level for the prediction error variance (PEV)
  # calculate the adjusted variance of BLUPs
  vblup <- 2 * mean(g_convar)
  # calculate the average standard error of differences (AVSED)
  avsed <- sqrt(vblup)
  H2 <- 1 - (vblup / (2 * var_g))
  out_list <- list("avsed" = avsed, "H2" = H2)
  return(out_list)}
```
      ## Heritability of PreGRArea
```{r}
dfq2.2<-df4[complete.cases(df4$PreGRArea),]
dfq2.2.0<-df4Mean[complete.cases(df4Mean$MPreGRArea),]

mod0<-glmmTMB(data=dfq2.2, PreGRArea~TreatEnv+(1|Genotype),family=Gamma(log),dispformula =~TreatEnv)
sim<-simulateResiduals(mod0,n=500, plot=T);testDispersion(sim)
summary(mod0)
car::Anova(mod0)
Cullis_H2TMB(mod0)

mod2.2<-glmmTMB(data=dfq2.2[dfq2.2$TreatEnv=="Control",], PreGRArea~1+(1|Genotype))
sim<-simulateResiduals(mod2.2,n=500, plot=T);testDispersion(sim)
summary(mod2)
Cullis_H2TMB(mod2.2)

mod1.2<-glmmTMB(data=dfq2.2[dfq2.2$TreatEnv=="Copper",], PreGRArea~1+(1|Genotype))
sim<-simulateResiduals(mod1.2,n=500, plot=T);testDispersion(sim)
summary(mod1)
Cullis_H2TMB(mod1.2)

rsq(mod)
rsq.partial(mod)
performance::r2(mod)
```


       ## calculate h2
```{r}
# Assume 'GRM' is a genomic relationship matrix created from SNP data
# and 'data' is your phenotype data with fitness trait and genotypes.
GRM<-read.table("./Kinship_IBD_matrix_42geno.txt", header = T, row.names = 1)
GRM <- as.matrix(GRM)
if (!isSymmetric(GRM)) {
  stop("The loaded matrix is not symmetric. Please check the input format.")}
# Confirm that the matrix dimensions are correct
print(dim(GRM))


### EXTRACT TABLE FROM TASSEL WITH CORRECT DATA

model <- mmer(PreGRArea ~ 1, 
              random = ~vsr(Genotype, Gu = GRM),# Specify GRM as random effect
              rcov=~units,
              data = dfq2.3[dfq2.3$TreatEnv=="Control",])  ##0.208258

model <- mmer(PreGRArea ~ 1, 
              random = ~vsr(Genotype, Gu = GRM),# Specify GRM as random effect
              rcov=~units,
              data = dfq2.3[dfq2.3$TreatEnv=="Copper",])  ##0.309337

# Extract variance components and calculate h2
var_components <- summary(model)$varcomp
V_A <- var_components["u:Genotype.PreGRArea-PreGRArea", "VarComp"]  # Genetic variance
V_E <- var_components["units.PreGRArea-PreGRArea", "VarComp"]       # Residual varianc
h2 <- V_A / (V_A + V_E)  # Narrow-sense heritability
h2

```


##################################################
##################################################
##################################################
###### Analysis of metabolites ###################
##################################################
##################################################
##################################################

#upload data of metabolites
```{r}
dfMet0<-read_excel("./GeneticBase_TransgnerationalPlasticity.xlsx", col_names = T, na = "",sheet="Metabolites")
```
  #convert to factors
```{r}
# create ordered variables for the plots
dfMet0$Pretreat1<-factor(dfMet0$Pretreat, levels=c("Cu-","Cu+"), ordered=TRUE)
dfMet0$TreatEnv1<-factor(dfMet0$TreatEnv, levels=c("Control","Copper"), ordered=TRUE) 

# define class of variables 
dfMet0$Pretreat<-factor(dfMet0$Pretreat)
dfMet0$TreatEnv<-factor(dfMet0$TreatEnv)
dfMet0$Genotype<-factor(dfMet0$Genotype)
dfMet0$Plate<-factor(dfMet0$Plate)
dfMet0$Column<-factor(dfMet0$Column)
dfMet0$SamplePosition<-factor(dfMet0$SamplePosition)
```
#### Calclulating the pre-treatment ratio for metabolites
```{r}
dfMet1<-data.table(dfMet0)
TreatEnv<-unique(dfMet1$TreatEnv)
Metabolite<-colnames(dfMet1)[7:12]

for (m in TreatEnv){
  for (n in Metabolite){ 
        a5<-as.numeric(with(dfMet1[TreatEnv==m&Pretreat=="Cu-",], mean(get(n),na.rm=T)))
                            dfMet1[TreatEnv==m&Pretreat=="Cu+",paste0("Pre",n):=get(n)/round(a5,8)]
  }}
```

 Prepare fitness data and merge with metabolite data
```{r}
df_summ<-df3 %>% 
  dplyr::group_by(Genotype,Pretreat, TreatEnv,Population) %>% dplyr::summarise(
  meanGRArea=mean(GRArea, na.rm = TRUE),
  meanPreGRArea=mean(PreGRArea, na.rm = TRUE),
  meanOffFvFm=mean(OffFvFm, na.rm = TRUE),
  meanPreOffFvFm=mean(PreOffFvFm, na.rm = TRUE))
df_summ<-data.table(df_summ)

df_Met<-merge(df_summ,dfMet1, by=c("Genotype","Pretreat", "TreatEnv"))
df_Met<-data.table(df_Met)
```

 Clean pre-treatment data
```{r}
df_Met2<-df_Met[Pretreat=="Cu+",]
df_Met2<-df_Met2[,-c(2,5,7,10,12:18)]

Metab<-colnames(df_Met2)[9:14]
dfsumm<-df_Met2 %>%
  dplyr::group_by(TreatEnv) %>%
  dplyr::summarise_at(vars(all_of(Metab)), mean, na.rm = TRUE)
```

 Mean of each genotype for the analysis and merge data
```{r}
df4Mean<-(df4 %>% dplyr::group_by(Genotype,TreatEnv,Population) %>% dplyr::summarise(
  MPreGRArea    = mean(PreGRArea, na.rm = TRUE)))
df4Mean<-data.table(df4Mean)

#To make even comparison with names (doesn't generate any mean)
dfMetMean<-(dfMet0 %>% dplyr::group_by(Genotype,Pretreat,TreatEnv) %>% dplyr::summarise(                
  MCyanidine3GlcHPLC    = mean(Cyanidine3GlcHPLC, na.rm = TRUE)   , 
  MCyanidinMalGlcHPLC   = mean(CyanidinMalGlcHPLC, na.rm = TRUE)  , 
    
  MLuteolin8CGlc  = mean(Luteolin8CGlc, na.rm = TRUE) , 
  MApigenin8CGlc = mean(Apigenin8CGlc, na.rm = TRUE), 

  MLuteolin7OGlc = mean(Luteolin7OGlc, na.rm = TRUE), 
  MApigenin7OGlc = mean(Apigenin7OGlc, na.rm = TRUE)
  ))

### For control plants against pre-treatment
dfMeanCtr<-dfMetMean[dfMetMean$TreatEnv=="Control"&dfMetMean$Pretreat=="Cu-",]
df4MeanCtr<-df4Mean[TreatEnv=="Control",]
df4MeanCu<-df4Mean[TreatEnv=="Copper",]

CtrPreCtr<-merge(dfMeanCtr,df4MeanCtr,by=c("Genotype"))
CtrPreCu<-merge(dfMeanCtr,df4MeanCu,by=c("Genotype"))
```

  ##Cyanidine3GlcHPLC
    #Pretreat*TreatEnv* 
```{r}
df1lm<-df_Met[complete.cases(df_Met$Cyanidine3GlcHPLC),]
options(contrasts = c("contr.treatment", "contr.poly"))

model1<-glmmTMB(data=df1lm, Cyanidine3GlcHPLC~Pretreat*TreatEnv+(1|Genotype),family=gaussian(inverse),dispformula=~TreatEnv)
sim1<-simulateResiduals(model1,n=500, plot=T);testDispersion(sim1)
summary(model1)

options(contrasts = c("contr.sum", "contr.poly"))
modelp<-glmmTMB(data=df1lm, Cyanidine3GlcHPLC~Pretreat*TreatEnv1+(1|Genotype),family=gaussian(inverse),dispformula=~TreatEnv)
car::Anova(modelp,type="III")

df1lm%>%dplyr::group_by(TreatEnv,Pretreat)%>%dplyr::summarise(data_mean=mean(Cyanidine3GlcHPLC, na.rm=T),N=length(Cyanidine3GlcHPLC))
((0.1147804-0.1137393	)/0.1137393	)*100 #0.9
((0.2365233-0.2221308)/0.2221308)*100 #6.5

moda<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Control",],Cyanidine3GlcHPLC~Pretreat+(1|Genotype))
moda<-glmmTMB(data=df1lm[df1lm$TreatEnv1=="Copper",],Cyanidine3GlcHPLC~Pretreat+(1|Genotype),family=gaussian(log))
simulateResiduals(moda,n=500, plot=T);#plotResiduals(model1,df1lm$Pretreat,main=NULL)
car::Anova(moda,type="II")

dflm1<-(df1lm %>% dplyr::group_by(TreatEnv1,Pretreat1) %>% dplyr::summarise(
  meanCyanidine3GlcHPLC   = mean(Cyanidine3GlcHPLC),
  se = (sd(Cyanidine3GlcHPLC)/sqrt(length(Cyanidine3GlcHPLC)))))

pdf("./FigS7b.pdf",height=6,width=4)
ggplot(df1lm,aes(Pretreat1, Cyanidine3GlcHPLC))+
  geom_col(data=dflm1,aes(x=Pretreat1, y=meanCyanidine3GlcHPLC,fill=Pretreat1),width=0.8)+
  geom_errorbar(data=dflm1,aes(x=Pretreat1, y=meanCyanidine3GlcHPLC,ymin=meanCyanidine3GlcHPLC-se, ymax=meanCyanidine3GlcHPLC+se), width=0.3, position=position_dodge(0.80))+
  scale_color_manual(values = c("Cu-"="#0056B3", "Cu+"="#CC0000"))+
  scale_fill_manual(values = c("Cu-"="#0056B3", "Cu+"="#CC0000"))+
  facet_grid(~TreatEnv1)+
  scale_y_continuous(breaks=seq(0,0.25,by=0.05),limits = c(0,0.25), expand = c(0,0))+
  guides(col="none",fill="none")+labs(y="Cyanidine3GlcHPLC µmol/g FW")+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
        axis.text.y = element_text(size = 25,color="black", family = "sans",angle=90),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        legend.text = element_text(size = 20,color="black", family = "sans"),
        legend.background = element_rect(fill = F),
        legend.key = element_rect(fill = F),
        legend.title = element_text(size = 30,color="black", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```
    #Induction # Cu-
```{r}
df1lm<-df_Met[complete.cases(df_Met$Cyanidine3GlcHPLC),]
df1lm<-df1lm[df1lm$Pretreat=="Cu-",]

model1<-glmmTMB(data=df1lm, Cyanidine3GlcHPLC~TreatEnv+(1|Genotype),family=Gamma(log))
plotResiduals(model1,df1lm$TreatEnv,main=NULL);sim1<-simulateResiduals(model1,n=500, plot=T);testDispersion(sim1)
model1<-glmmTMB(data=df1lm, Cyanidine3GlcHPLC~TreatEnv+(1|Genotype),family=Gamma(log),contrasts=list(TreatEnv="contr.sum"))
summary(model1)
car::Anova(model1,type="II")

dflm1<-(df1lm %>% dplyr::group_by(TreatEnv1) %>% dplyr::summarise(
  meanCyanidine3GlcHPLC   = mean(Cyanidine3GlcHPLC),
  se = (sd(Cyanidine3GlcHPLC)/sqrt(length(Cyanidine3GlcHPLC)))))
((0.008735452-0.002707439)/0.002707439)*100 ##222.6

pdf("./FigS7a2.pdf",height=6,width=4)
ggplot(df1lm,aes(TreatEnv1, Cyanidine3GlcHPLC))+
  geom_col(data=dflm1,aes(x=TreatEnv1, y=meanCyanidine3GlcHPLC,fill=TreatEnv1),width=0.8)+
  geom_errorbar(data=dflm1,aes(x=TreatEnv1, y=meanCyanidine3GlcHPLC,ymin=meanCyanidine3GlcHPLC-se, ymax=meanCyanidine3GlcHPLC+se), width=0.3, position=position_dodge(0.80))+
  scale_color_manual(values = c("Control"="#0056B3", "Copper"="#CC0000"))+
  scale_fill_manual(values = c("Control"="#0056B3", "Copper"="#CC0000"))+
  scale_y_continuous(breaks=seq(0.01,0.25,by=0.08),limits = c(0,0.25), expand = c(0,0))+
  guides(col="none",fill="none")+labs(y="Cyanidine3GlcHPLC µmol/g FW")+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
        axis.text.y = element_text(size = 25,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```
     #Correlation with mean population growth rate in naive plants
```{r}
df1lm<-df1lm[df1lm$Pretreat=="Cu-",]
options(contrasts = c("contr.treatment", "contr.poly"))
model<-glmmTMB(data=df1lm, meanGRArea~Cyanidine3GlcHPLC*TreatEnv+(1|Plate))
sim<-simulateResiduals(model,n=500, plot=T)
summary(model)
options(contrasts = c("contr.sum", "contr.poly"))
model<-glmmTMB(data=df1lm, meanGRArea~Cyanidine3GlcHPLC*TreatEnv+(1|Plate))
car::Anova(model,type=3)

model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Control",], meanGRArea~Cyanidine3GlcHPLC+(1|Plate))
model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Copper",], meanGRArea~Cyanidine3GlcHPLC+(1|Plate))
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model,type=2)
plot(allEffects(model))

pdf("./FigS5b.pdf",height=6,width=8)
ggplot(df1lm,aes(y=meanGRArea, x=Cyanidine3GlcHPLC,col=TreatEnv))+
  geom_point(size=3)+
  scale_color_manual(values = c("Control"="#0056B3", "Copper"="#CC0000"))+
  geom_smooth(method="lm",se=F,size=2)+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5, npc.x=1.8)+
  scale_y_continuous(breaks=seq(0.2,0.5,by=0.1),limits = c(0.2,0.5), expand = c(0,0))+
  scale_x_continuous(breaks=seq(0,0.5,by=0.1),limits = c(0,0.5), expand = c(0,0))+
  guides(col="none")+
  theme(axis.text.x = element_text(size = 20,color="black", family = "sans"),
        axis.text.y = element_text(size = 20,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```
     #Correlation with mean pre-treatment ratio of population growth rate
```{r}
df1lm<-df_Met2[complete.cases(df_Met2$PreCyanidine3GlcHPLC),]
options(contrasts = c("contr.treatment", "contr.poly"))

model<-glmmTMB(data=df1lm, meanPreGRArea~(PreCyanidine3GlcHPLC)*TreatEnv+(1|Genotype),family=Gamma(log))
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model)

pdf("./FigS7d1.pdf",height=6,width=5)
ggplot(df1lm,aes(y=meanPreGRArea, x=PreCyanidine3GlcHPLC,col=TreatEnv))+
  geom_point(size=3)+
  scale_color_manual(values = c("Control"="#0056B3", "Copper"="#CC0000"))+
  geom_smooth(method="lm",se=F,size=3)+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  scale_y_continuous(breaks=seq(0.7,1.5,by=0.2),limits = c(0.7,1.5), expand = c(0,0))+
  scale_x_continuous(breaks=seq(0.5,2.5,by=0.5),limits = c(0.5,2.5), expand = c(0,0))+
  guides(col="none")+
  theme(axis.text.x = element_text(size = 20,color="black", family = "sans"),
        axis.text.y = element_text(size = 20,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```
     #Correlation with mean FvFm
```{r}
df1lm<-df_Met[complete.cases(df_Met$Cyanidine3GlcHPLC),]
df1lm<-df1lm[complete.cases(df1lm$meanOffFvFm),]
df1lm<-df1lm[df1lm$Pretreat=="Cu-",]
options(contrasts = c("contr.treatment", "contr.poly"))

model<-glmmTMB(data=df1lm, meanOffFvFm~Cyanidine3GlcHPLC*TreatEnv+(1|Plate),dispformula = ~TreatEnv)
sim<-simulateResiduals(model,n=500, plot=T)
summary(model)
model<-glmmTMB(data=df1lm, meanOffFvFm~Cyanidine3GlcHPLC*TreatEnv+(1|Plate),dispformula = ~TreatEnv,contrasts=list(TreatEnv="contr.sum"))
car::Anova(model,type=3)

model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Control"], meanOffFvFm~Cyanidine3GlcHPLC+(1|Plate))
model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Copper"], meanOffFvFm~Cyanidine3GlcHPLC+(1|Plate),family=gaussian(log))
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model,type=2)

pdf("./FigS4a.pdf",height=6,width=8)
ggplot(df1lm,aes(y=meanOffFvFm, x=Cyanidine3GlcHPLC,col=TreatEnv))+
  geom_point(size=3)+
  scale_color_manual(values = c("Control"="#0056B3", "Copper"="#CC0000"))+
  geom_smooth(method="lm",se=F,size=2)+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5, npc.x=1.8)+
  scale_y_continuous(breaks=seq(0.6,0.8,by=0.05),limits = c(0.65,0.8), expand = c(0,0))+
  scale_x_continuous(breaks=seq(0,0.5,by=0.1),limits = c(0,0.5), expand = c(0,0))+
  guides(col="none")+
  theme(axis.text.x = element_text(size = 20,color="black", family = "sans"),
        axis.text.y = element_text(size = 20,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        legend.text = element_text(size = 20,color="black", family = "sans"),
        legend.background = element_rect(fill = F),
        legend.key = element_rect(fill = F),
        legend.title = element_text(size = 30,color="black", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```
     #Correlation with mean pre-treatment ratio of FvFm
```{r}
df1lm<-df_Met2[complete.cases(df_Met2$PreCyanidine3GlcHPLC),]
df1lm<-df1lm[complete.cases(df1lm$meanPreOffFvFm),]
options(contrasts = c("contr.treatment", "contr.poly"))

model<-glmmTMB(data=df1lm, meanPreOffFvFm~(PreCyanidine3GlcHPLC)*TreatEnv)
sim<-simulateResiduals(model,n=500, plot=T)
summary(model)
model<-glmmTMB(data=df1lm, meanPreOffFvFm~(PreCyanidine3GlcHPLC)*TreatEnv,contrasts=list(TreatEnv="contr.sum"))
car::Anova(model,type=3)
plot(allEffects(model))

pdf("./FigS7c1.pdf",height=6,width=5)
ggplot(df1lm,aes(y=meanPreOffFvFm, x=PreCyanidine3GlcHPLC,col=TreatEnv))+
  geom_point(size=3)+
  scale_color_manual(values = c("Control"="#0056B3", "Copper"="#CC0000"))+
  geom_smooth(method="lm",se=F,size=3)+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  scale_y_continuous(breaks=seq(0.85,1.15,by=0.1),limits = c(0.85,1.15), expand = c(0,0))+
  scale_x_continuous(breaks=seq(0.5,2.5,by=0.5),limits = c(0.5,2.5), expand = c(0,0))+
  guides(col="none")+
  theme(axis.text.x = element_text(size = 20,color="black", family = "sans"),
        axis.text.y = element_text(size = 20,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```
       # correlation between growth rate under control and plant pre-treatment
```{r}
model<-glmmTMB(data=CtrPreCtr,MCyanidine3GlcHPLC~MPreGRArea)
plotResiduals(model,CtrPreCtr$MCyanidine3GlcHPLC);plotResiduals(model,CtrPreCtr$MPreGRArea)
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model)

pdf("./FigS4a1.pdf",height=6,width=6)
ggplot(data=CtrPreCtr,aes(x=MPreGRArea,y=MCyanidine3GlcHPLC))+
  geom_point(size=3)+
  labs(y="Cyanidine3GlcHPLC under control",x="Pre-treatment ratio under control")+
  geom_line(aes(y=predict(model, type="response")),size=1.5,col="#CC0000")+
  scale_x_continuous(breaks=seq(0.9,1.2,by=0.1),limits=c(0.9,1.2),expand=c(0,0))+
  scale_y_continuous(breaks=seq(0.08,0.21,by=0.02),limits=c(0.08,0.21),expand=c(0,0))+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()


model<-glmmTMB(data=CtrPreCu,MCyanidine3GlcHPLC~MPreGRArea)
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model)

pdf("./FigS4a2.pdf",height=6,width=6)
ggplot(data=CtrPreCu,aes(x=MPreGRArea,y=MCyanidine3GlcHPLC))+
  geom_point(size=3)+
  labs(y="Cyanidine3GlcHPLC under control",x="Pre-treatment ratio under copper")+
  geom_line(aes(y=predict(model, type="response")),size=1.5,col="#CC0000")+
  scale_x_continuous(breaks=seq(0.7,1.5,by=0.2),limits=c(0.7,1.5),expand=c(0,0))+
  scale_y_continuous(breaks=seq(0.08,0.21,by=0.02),limits=c(0.08,0.21),expand=c(0,0))+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()
```


  ##CyanidinMalGlcHPLC
    #Pretreat*TreatEnv* 
```{r}
df1lm<-df_Met[complete.cases(df_Met$CyanidinMalGlcHPLC),]
options(contrasts = c("contr.treatment", "contr.poly"))

model1<-glmmTMB(data=df1lm, CyanidinMalGlcHPLC~Pretreat*TreatEnv+(1|Genotype),family=Gamma(log))
plotResiduals(model1,df1lm$TreatEnv,main=NULL);plotResiduals(model1,df1lm$Pretreat,main=NULL);
sim1<-simulateResiduals(model1,n=500, plot=T);testDispersion(sim1)
model1<-glmmTMB(data=df1lm, CyanidinMalGlcHPLC~Pretreat*TreatEnv1+(1|Genotype),family=Gamma(log),contrasts=list(Pretreat="contr.sum",TreatEnv1="contr.sum"))
summary(model1)
car::Anova(model1,type="III")

df1lm%>%dplyr::group_by(TreatEnv,Pretreat)%>%dplyr::summarise(data_mean=mean(CyanidinMalGlcHPLC, na.rm=T),N=length(CyanidinMalGlcHPLC))
((0.3151211-0.2999643)/0.2999643)*100 #5
((0.7969298-0.7586004)/0.7586004)*100 #5

moda<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Control",],CyanidinMalGlcHPLC~Pretreat+(1|Genotype),family=gaussian(log))
moda<-glmmTMB(data=df1lm[df1lm$TreatEnv1=="Copper",],CyanidinMalGlcHPLC~Pretreat+(1|Genotype),family=gaussian(log))
simulateResiduals(moda,n=500, plot=T);#plotResiduals(model1,df1lm$Pretreat,main=NULL)
car::Anova(moda,type="II")

dflm1<-(df1lm %>% dplyr::group_by(TreatEnv1,Pretreat1) %>% dplyr::summarise(
  meanCyanidinMalGlcHPLC   = mean(CyanidinMalGlcHPLC),
  se = (sd(CyanidinMalGlcHPLC)/sqrt(length(CyanidinMalGlcHPLC)))))

pdf("./Fig4d.pdf",height=6,width=4)
ggplot(df1lm,aes(Pretreat1, CyanidinMalGlcHPLC))+
  geom_col(data=dflm1,aes(x=Pretreat1, y=meanCyanidinMalGlcHPLC,fill=Pretreat1),width=0.8)+
  geom_errorbar(data=dflm1,aes(x=Pretreat1, y=meanCyanidinMalGlcHPLC,ymin=meanCyanidinMalGlcHPLC-se, ymax=meanCyanidinMalGlcHPLC+se), width=0.3, position=position_dodge(0.80))+
  scale_color_manual(values = c("Cu-"="#0056B3", "Cu+"="#CC0000"))+
  scale_fill_manual(values = c("Cu-"="#0056B3", "Cu+"="#CC0000"))+
  facet_grid(~TreatEnv1)+
  scale_y_continuous(breaks=seq(0.1,0.9,by=0.2),limits = c(0,0.9), expand = c(0,0))+
  guides(col="none",fill="none")+labs(y="CyanidinMalGlcHPLC µmol/g FW")+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
        axis.text.y = element_text(size = 25,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        legend.text = element_text(size = 20,color="black", family = "sans"),
        legend.background = element_rect(fill = F),
        legend.key = element_rect(fill = F),
        legend.title = element_text(size = 30,color="black", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```
    #Induction # Cu-
```{r}
df1lm<-df_Met[complete.cases(df_Met$CyanidinMalGlcHPLC),]
df1lm<-df1lm[df1lm$Pretreat=="Cu-",]

model1<-glmmTMB(data=df1lm, CyanidinMalGlcHPLC~TreatEnv+(1|Genotype),family=Gamma(log))
plotResiduals(model1,df1lm$TreatEnv,main=NULL);sim1<-simulateResiduals(model1,n=500, plot=T);testDispersion(sim1)
model1<-glmmTMB(data=df1lm, CyanidinMalGlcHPLC~TreatEnv+(1|Genotype),family=Gamma(log),contrasts=list(TreatEnv="contr.sum"))
summary(model1)
car::Anova(model1,type="II")

dflm1<-(df1lm %>% dplyr::group_by(TreatEnv1) %>% dplyr::summarise(
  meanCyanidinMalGlcHPLC   = mean(CyanidinMalGlcHPLC),
  se = (sd(CyanidinMalGlcHPLC)/sqrt(length(CyanidinMalGlcHPLC)))))
((0.7586004-0.2999643)/0.2999643)*100 #152.8969

pdf("./FigS7a1.pdf",height=6,width=4)
ggplot(df1lm,aes(TreatEnv1, CyanidinMalGlcHPLC))+
  geom_col(data=dflm1,aes(x=TreatEnv1, y=meanCyanidinMalGlcHPLC,fill=TreatEnv1),width=0.8)+
  geom_errorbar(data=dflm1,aes(x=TreatEnv1, y=meanCyanidinMalGlcHPLC,ymin=meanCyanidinMalGlcHPLC-se, ymax=meanCyanidinMalGlcHPLC+se), width=0.3, position=position_dodge(0.80))+
  scale_color_manual(values = c("Control"="#0056B3", "Copper"="#CC0000"))+
  scale_fill_manual(values = c("Control"="#0056B3", "Copper"="#CC0000"))+
  scale_y_continuous(breaks=seq(0,0.8,by=0.2),limits = c(0,0.8), expand = c(0,0))+
  guides(col="none",fill="none")+labs(y="CyanidinMalGl µmol/g FW")+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
        axis.text.y = element_text(size = 25,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```
     #Correlation with mean population growth rate in naive plants
```{r}
df1lm<-df1lm[df1lm$Pretreat=="Cu-",]
model<-glmmTMB(data=df1lm, meanGRArea~CyanidinMalGlcHPLC*TreatEnv+(1|Plate))
sim<-simulateResiduals(model,n=500, plot=T)
model<-glmmTMB(data=df1lm, meanGRArea~CyanidinMalGlcHPLC*TreatEnv+(1|Plate),contrasts=list(TreatEnv="contr.sum"))
car::Anova(model,type=3)
summary(model)
model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Control",], meanGRArea~CyanidinMalGlcHPLC+(1|Plate))
model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Copper",], meanGRArea~CyanidinMalGlcHPLC+(1|Plate))
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model,type=2)
plot(allEffects(model))

pdf("./Fig4c.pdf",height=6,width=8)
ggplot(df1lm,aes(y=meanGRArea, x=CyanidinMalGlcHPLC,col=TreatEnv))+
  geom_point(size=3)+
  scale_color_manual(values = c("Control"="#0056B3", "Copper"="#CC0000"))+
  geom_smooth(method="lm",se=F,size=2)+
  stat_poly_eq(formula = x~y, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  scale_y_continuous(breaks=seq(0.2,0.5,by=0.1),limits = c(0.2,0.5), expand = c(0,0))+
  scale_x_continuous(breaks=seq(0,2,by=0.5),limits = c(0,2), expand = c(0,0))+
  guides(col="none")+
  theme(axis.text.x = element_text(size = 20,color="black", family = "sans"),
        axis.text.y = element_text(size = 20,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        legend.text = element_text(size = 20,color="black", family = "sans"),
        legend.background = element_rect(fill = F),
        legend.key = element_rect(fill = F),
        legend.title = element_text(size = 30,color="black", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```
     #Correlation with mean pre-treatment ratio of population growth rate
```{r}
df1lm<-df_Met2[complete.cases(df_Met2$PreCyanidinMalGlcHPLC),]
df1lm<-df1lm[complete.cases(df1lm$meanPreGRArea),]
options(contrasts = c("contr.treatment", "contr.poly"))

model<-glmmTMB(data=df1lm, meanPreGRArea~PreCyanidinMalGlcHPLC*TreatEnv+(1|Genotype),family=Gamma(log))
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model)

pdf("./FigS7d2.pdf",height=6,width=5)
ggplot(df1lm,aes(y=meanPreGRArea, x=PreCyanidinMalGlcHPLC,col=TreatEnv))+
  geom_point(size=3)+
  scale_color_manual(values = c("Control"="#0056B3", "Copper"="#CC0000"))+
  geom_smooth(method="lm",se=F,size=3)+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  scale_y_continuous(breaks=seq(0.7,1.5,by=0.2),limits = c(0.7,1.5), expand = c(0,0))+
  scale_x_continuous(breaks=seq(0.4,2.4,by=0.5),limits = c(0.4,2.4), expand = c(0,0))+
  guides(col="none")+
  theme(axis.text.x = element_text(size = 20,color="black", family = "sans"),
        axis.text.y = element_text(size = 20,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```
     #Correlation with mean FvFm
```{r}
df1lm<-df_Met[complete.cases(df_Met$CyanidinMalGlcHPLC),]
df1lm<-df1lm[complete.cases(df1lm$meanOffFvFm),]
df1lm<-df1lm[df1lm$Pretreat=="Cu-",]
options(contrasts = c("contr.treatment", "contr.poly"))

model<-glmmTMB(data=df1lm, meanOffFvFm~CyanidinMalGlcHPLC*TreatEnv+(1|Plate),family=Gamma(log))
model<-glmmTMB(data=df1lm, meanOffFvFm~CyanidinMalGlcHPLC*TreatEnv+(1|Plate),dispformula = ~TreatEnv)
sim<-simulateResiduals(model,n=500, plot=T)
summary(model)
options(contrasts = c("contr.sum", "contr.poly"))
modelp<-glmmTMB(data=df1lm, meanOffFvFm~CyanidinMalGlcHPLC*TreatEnv+(1|Plate),dispformula = ~TreatEnv)
car::Anova(modelp,type=3)
plot(allEffects(model))

model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Control",], meanOffFvFm~CyanidinMalGlcHPLC+(1|Plate))
model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Copper",], meanOffFvFm~CyanidinMalGlcHPLC+(1|Plate))
sim<-simulateResiduals(model,n=500, plot=T)
summary(model)
car::Anova(model)

pdf("./Fig3b.pdf",height=6,width=8)
ggplot(df1lm,aes(y=meanOffFvFm, x=CyanidinMalGlcHPLC,col=TreatEnv))+
  geom_point(size=3)+
  scale_color_manual(values = c("Control"="#0056B3", "Copper"="#CC0000"))+
  geom_smooth(method="lm",se=F,size=2)+
  stat_poly_eq(formula = x~y, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  scale_y_continuous(breaks=seq(0.6,0.8,by=0.05),limits = c(0.65,0.8), expand = c(0,0))+
  scale_x_continuous(breaks=seq(0,2,by=0.5),limits = c(0,2), expand = c(0,0))+
  guides(col="none")+
  theme(axis.text.x = element_text(size = 20,color="black", family = "sans"),
        axis.text.y = element_text(size = 20,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        legend.text = element_text(size = 20,color="black", family = "sans"),
        legend.background = element_rect(fill = F),
        legend.key = element_rect(fill = F),
        legend.title = element_text(size = 30,color="black", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```
     #Correlation with mean pre-treatment ratio of FvFm
```{r}
df1lm<-df_Met2[complete.cases(df_Met2$PreCyanidinMalGlcHPLC),]
df1lm<-df1lm[complete.cases(df1lm$meanPreOffFvFm),]
options(contrasts = c("contr.treatment", "contr.poly"))

model<-glmmTMB(data=df1lm, PreCyanidinMalGlcHPLC~meanPreOffFvFm*TreatEnv,family=Gamma(log))
sim<-simulateResiduals(model,n=500, plot=T)
summary(model)
model<-glmmTMB(data=df1lm, meanPreOffFvFm~PreCyanidinMalGlcHPLC*TreatEnv,contrasts=list(TreatEnv="contr.sum"))
model<-glmmTMB(data=df1lm, PreCyanidinMalGlcHPLC~meanPreOffFvFm*TreatEnv,family=Gamma(log),contrasts=list(TreatEnv="contr.sum"))
car::Anova(model,type=3)
plot(allEffects(model))

pdf("./FigS7c2.pdf",height=6,width=5)
ggplot(df1lm,aes(y=meanPreOffFvFm, x=PreCyanidinMalGlcHPLC,col=TreatEnv))+
  geom_point(size=3)+
  scale_color_manual(values = c("Control"="#0056B3", "Copper"="#CC0000"))+
  geom_smooth(method="lm",se=F,size=3)+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  scale_y_continuous(breaks=seq(0.85,1.15,by=0.1),limits = c(0.85,1.15), expand = c(0,0))+
  scale_x_continuous(breaks=seq(0.4,2.4,by=0.5),limits = c(0.4,2.4), expand = c(0,0))+
  guides(col="none")+
  theme(axis.text.x = element_text(size = 20,color="black", family = "sans"),
        axis.text.y = element_text(size = 20,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```
       # correlation between growth rate under control and plant pre-treatment
```{r}
model<-glmmTMB(data=CtrPreCtr,MCyanidinMalGlcHPLC~MPreGRArea)
#model<-glmmTMB(data=CtrPreCtr,MCyanidinMalGlcHPLC~MPreGRArea+(1|Population))
plotResiduals(model,CtrPreCtr$MCyanidinMalGlcHPLC);plotResiduals(model,CtrPreCtr$MPreGRArea)
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model)

pdf("./FigS4a3.pdf",height=6,width=6)
ggplot(data=CtrPreCtr,aes(x=MPreGRArea,y=MCyanidinMalGlcHPLC))+
  geom_point(size=3)+
  labs(y="CyanidinMalGlcHPLC under control",x="Pre-treatment ratio under control")+
  geom_line(aes(y=predict(model, type="response")),size=1.5,col="#CC0000")+
  scale_x_continuous(breaks=seq(0.9,1.2,by=0.1),limits=c(0.9,1.2),expand=c(0,0))+
  scale_y_continuous(breaks=seq(0.1,0.7,by=0.2),limits=c(0.1,0.7),expand=c(0,0))+
  #stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()


model<-glmmTMB(data=CtrPreCu,MCyanidinMalGlcHPLC~MPreGRArea)
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model)

pdf("./FigS4a4.pdf",height=6,width=6)
ggplot(data=CtrPreCu,aes(x=MPreGRArea,y=MCyanidinMalGlcHPLC))+
  geom_point(size=3)+
  labs(y="CyanidinMalGlcHPLC under control",x="Pre-treatment ratio under copper")+
  geom_line(aes(y=predict(model, type="response")),size=1.5,col="#CC0000")+
  scale_x_continuous(breaks=seq(0.7,1.5,by=0.2),limits=c(0.7,1.5),expand=c(0,0))+
  scale_y_continuous(breaks=seq(0.1,0.7,by=0.2),limits=c(0.1,0.7),expand=c(0,0))+
  #stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()
```


  ##Luteolin7OGlc
  #Correlation with mean population growth rate in naive plants
```{r}
df1lm<-df_Met[complete.cases(df_Met$Luteolin7OGlc),]
df1lm<-df1lm[df1lm$Pretreat=="Cu-",]
options(contrasts = c("contr.treatment", "contr.poly"))

model<-glmmTMB(data=df1lm, meanGRArea~Luteolin7OGlc*TreatEnv+(1|Plate))
sim<-simulateResiduals(model,n=500, plot=T)
model<-glmmTMB(data=df1lm, meanGRArea~Luteolin7OGlc*TreatEnv+(1|Plate),contrasts=list(TreatEnv="contr.sum"))
car::Anova(model,type=3)
summary(model)
model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Control",], meanGRArea~Luteolin7OGlc+(1|Plate))
model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Copper",], meanGRArea~Luteolin7OGlc+(1|Plate))
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model,type=2)
plot(allEffects(model))

pdf("./FigS6b1.pdf",height=6,width=6)
ggplot(df1lm,aes(y=meanGRArea, x=Luteolin7OGlc,col=TreatEnv))+
  geom_point(size=3)+
  scale_color_manual(values = c("Control"="#0056B3", "Copper"="#CC0000"))+
  geom_smooth(method="lm",se=F,size=3)+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  scale_y_continuous(breaks=seq(0.2,0.5,by=0.1),limits = c(0.2,0.5), expand = c(0,0))+
  scale_x_continuous(breaks=seq(0,6,by=2),limits = c(0,6), expand = c(0,0))+
  guides(col="none")+
  theme(axis.text.x = element_text(size = 20,color="black", family = "sans"),
        axis.text.y = element_text(size = 20,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```
    #Correlation with mean FvFm in naive plants
```{r}
df1lm<-df_Met[complete.cases(df_Met$Luteolin7OGlc),]
df1lm<-df1lm[complete.cases(df1lm$meanOffFvFm),]
df1lm<-df1lm[df1lm$Pretreat=="Cu-",]
options(contrasts = c("contr.treatment", "contr.poly"))

model<-glmmTMB(data=df1lm, meanOffFvFm~Luteolin7OGlc*TreatEnv+(1|Plate),dispformula = ~TreatEnv)
sim<-simulateResiduals(model,n=500, plot=T)
summary(model)
options(contrasts = c("contr.sum", "contr.poly"))
modelp<-glmmTMB(data=df1lm, meanOffFvFm~Luteolin7OGlc*TreatEnv+(1|Plate),dispformula = ~TreatEnv)
car::Anova(modelp,type=3)
plot(allEffects(model))

model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Control",], meanOffFvFm~Luteolin7OGlc)
model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Copper",], meanOffFvFm~Luteolin7OGlc)
sim<-simulateResiduals(model,n=500, plot=T)
summary(model)
car::Anova(model)

pdf("./FigS6a1.pdf",height=6,width=6)
ggplot(df1lm,aes(y=meanOffFvFm, x=Luteolin7OGlc,col=TreatEnv))+
  geom_point(size=3)+
  scale_color_manual(values = c("Control"="#0056B3", "Copper"="#CC0000"))+
  geom_smooth(method="lm",se=F,size=2)+
  stat_poly_eq(formula = x~y, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  scale_y_continuous(breaks=seq(0.65,0.8,by=0.05),limits = c(0.65,0.8), expand = c(0,0))+
  scale_x_continuous(breaks=seq(0,6,by=2),limits = c(0,6), expand = c(0,0))+
  guides(col="none")+
  theme(axis.text.x = element_text(size = 20,color="black", family = "sans"),
        axis.text.y = element_text(size = 20,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```
       # correlation between growth rate under control and plant pre-treatment
```{r}
model<-glmmTMB(data=CtrPreCtr,MLuteolin7OGlc~MPreGRArea)
plotResiduals(model,CtrPreCtr$MLuteolin7OGlc);plotResiduals(model,CtrPreCtr$MPreGRArea)
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model)

pdf("./FigS4b1.pdf",height=6,width=6)
ggplot(data=CtrPreCtr,aes(x=MPreGRArea,y=MLuteolin7OGlc))+
  geom_point(size=3)+
  labs(y="Luteolin7OGlc under control",x="Pre-treatment ratio under control")+
  geom_line(aes(y=predict(model, type="response")),size=1.5,col="#CC0000")+
  scale_x_continuous(breaks=seq(0.9,1.2,by=0.1),limits=c(0.9,1.2),expand=c(0,0))+
  scale_y_continuous(breaks=seq(0,3,by=1),limits=c(0,3),expand=c(0,0))+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()

model<-glmmTMB(data=CtrPreCu,MLuteolin7OGlc~MPreGRArea)
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model)

pdf("./FigS4b2.pdf",height=6,width=6)
ggplot(data=CtrPreCu,aes(x=MPreGRArea,y=MLuteolin7OGlc))+
  geom_point(size=3)+
  labs(y="Luteolin7OGlc under control",x="Pre-treatment ratio under copper")+
  geom_line(aes(y=predict(model, type="response")),size=1.5,col="#CC0000")+
  scale_x_continuous(breaks=seq(0.7,1.5,by=0.2),limits=c(0.7,1.5),expand=c(0,0))+
  scale_y_continuous(breaks=seq(0,3,by=1),limits=c(0,3),expand=c(0,0))+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()
```

  ##Luteolin8CGlc
      #Correlation with mean population growth rate in naive plants
```{r}
df1lm<-df_Met[complete.cases(df_Met$Luteolin8CGlc),]
df1lm<-df1lm[df1lm$Pretreat=="Cu-",]
options(contrasts = c("contr.treatment", "contr.poly"))

model<-glmmTMB(data=df1lm, meanGRArea~Luteolin8CGlc*TreatEnv+(1|Plate))
sim<-simulateResiduals(model,n=500, plot=T)
model<-glmmTMB(data=df1lm, meanGRArea~Luteolin8CGlc*TreatEnv+(1|Plate),contrasts=list(TreatEnv="contr.sum"))
car::Anova(model,type=3)
summary(model)
model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Control",], meanGRArea~Luteolin8CGlc+(1|Plate))
model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Copper",], meanGRArea~Luteolin8CGlc+(1|Plate))
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model,type=2)
plot(allEffects(model))

pdf("./FigS6b2.pdf",height=6,width=6)
ggplot(df1lm,aes(y=meanGRArea, x=Luteolin8CGlc,col=TreatEnv))+
  geom_point(size=3)+
  scale_color_manual(values = c("Control"="#0056B3", "Copper"="#CC0000"))+
  geom_smooth(method="lm",se=F,size=3)+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  scale_y_continuous(breaks=seq(0.2,0.5,by=0.1),limits = c(0.2,0.5), expand = c(0,0))+
  scale_x_continuous(breaks=seq(0,4,by=1),limits = c(0,4), expand = c(0,0))+
  guides(col="none")+
  theme(axis.text.x = element_text(size = 20,color="black", family = "sans"),
        axis.text.y = element_text(size = 20,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```
    #Correlation with mean FvFm in naive plants
```{r}
df1lm<-df_Met[complete.cases(df_Met$Luteolin8CGlc),]
df1lm<-df1lm[complete.cases(df1lm$meanOffFvFm),]
df1lm<-df1lm[df1lm$Pretreat=="Cu-",]
options(contrasts = c("contr.treatment", "contr.poly"))

model<-glmmTMB(data=df1lm, meanOffFvFm~Luteolin8CGlc*TreatEnv+(1|Plate),dispformula = ~TreatEnv)
sim<-simulateResiduals(model,n=500, plot=T)
summary(model)
options(contrasts = c("contr.sum", "contr.poly"))
modelp<-glmmTMB(data=df1lm, meanOffFvFm~Luteolin8CGlc*TreatEnv+(1|Plate),dispformula = ~TreatEnv)
car::Anova(modelp,type=3)
plot(allEffects(model))

model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Control",], meanOffFvFm~Luteolin8CGlc)
model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Copper",], meanOffFvFm~Luteolin8CGlc)
sim<-simulateResiduals(model,n=500, plot=T)
summary(model)
car::Anova(model)

pdf("./FigS6a2.pdf",height=6,width=6)
ggplot(df1lm,aes(y=meanOffFvFm, x=Luteolin8CGlc,col=TreatEnv))+
  geom_point(size=3)+
  scale_color_manual(values = c("Control"="#0056B3", "Copper"="#CC0000"))+
  geom_smooth(method="lm",se=F,size=2)+
  stat_poly_eq(formula = x~y, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  scale_y_continuous(breaks=seq(0.65,0.8,by=0.05),limits = c(0.65,0.8), expand = c(0,0))+
  scale_x_continuous(breaks=seq(0,4,by=1),limits = c(0,4), expand = c(0,0))+
  guides(col="none")+
  theme(axis.text.x = element_text(size = 20,color="black", family = "sans"),
        axis.text.y = element_text(size = 20,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```
       # correlation between growth rate under control and plant pre-treatment
```{r}
model<-glmmTMB(data=CtrPreCtr,MLuteolin8CGlc~MPreGRArea)
plotResiduals(model,CtrPreCtr$MLuteolin8CGlc);plotResiduals(model,CtrPreCtr$MPreGRArea)
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model)

pdf("./FigS4b3.pdf",height=6,width=6)
ggplot(data=CtrPreCtr,aes(x=MPreGRArea,y=MLuteolin8CGlc))+
  geom_point(size=3)+
  labs(y="Luteolin8CGlc under control",x="Pre-treatment ratio under control")+
  geom_line(aes(y=predict(model, type="response")),size=1.5,col="#CC0000")+
  scale_x_continuous(breaks=seq(0.9,1.2,by=0.1),limits=c(0.9,1.2),expand=c(0,0))+
  scale_y_continuous(breaks=seq(0.5,2.5,by=0.5),limits=c(0.5,2.5),expand=c(0,0))+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()


model<-glmmTMB(data=CtrPreCu,MLuteolin8CGlc~MPreGRArea)
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model)

pdf("./FigS4b4.pdf",height=6,width=6)
ggplot(data=CtrPreCu,aes(x=MPreGRArea,y=MLuteolin8CGlc))+
  geom_point(size=3)+
  labs(y="Luteolin8CGlc under control",x="Pre-treatment ratio under copper")+
  geom_line(aes(y=predict(model, type="response")),size=1.5,col="#CC0000")+
  scale_x_continuous(breaks=seq(0.7,1.5,by=0.2),limits=c(0.7,1.5),expand=c(0,0))+
  scale_y_continuous(breaks=seq(0.5,2.5,by=0.5),limits=c(0.5,2.5),expand=c(0,0))+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()
```

  ##Apigenin7OGlc
     #Correlation with mean population growth rate in naive plants
```{r}
df1lm<-df_Met[complete.cases(df_Met$Apigenin7OGlc),]
df1lm<-df1lm[df1lm$Pretreat=="Cu-",]

options(contrasts = c("contr.treatment", "contr.poly"))
model<-glmmTMB(data=df1lm, meanGRArea~Apigenin7OGlc*TreatEnv+(1|Plate))
sim<-simulateResiduals(model,n=500, plot=T)
model<-glmmTMB(data=df1lm, meanGRArea~Apigenin7OGlc*TreatEnv+(1|Plate),contrasts=list(TreatEnv="contr.sum"))
car::Anova(model,type=3)
summary(model)
model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Control",], meanGRArea~Apigenin7OGlc+(1|Plate))
model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Copper",], meanGRArea~Apigenin7OGlc+(1|Plate))
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model,type=2)
plot(allEffects(model))

pdf("./FigS6b3.pdf",height=6,width=6)
ggplot(df1lm,aes(y=meanGRArea, x=Apigenin7OGlc,col=TreatEnv))+
  geom_point(size=3)+
  scale_color_manual(values = c("Control"="#0056B3", "Copper"="#CC0000"))+
  geom_smooth(method="lm",se=F,size=3)+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  scale_y_continuous(breaks=seq(0.2,0.5,by=0.1),limits = c(0.2,0.5), expand = c(0,0))+
  scale_x_continuous(breaks=seq(0,2,by=0.5),limits = c(0,2), expand = c(0,0))+
  guides(col="none")+
  theme(axis.text.x = element_text(size = 20,color="black", family = "sans"),
        axis.text.y = element_text(size = 20,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```
    #Correlation with mean FvFm in in naive plants
```{r}
df1lm<-df_Met[complete.cases(df_Met$Apigenin7OGlc),]
df1lm<-df1lm[complete.cases(df1lm$meanOffFvFm),]
df1lm<-df1lm[df1lm$Pretreat=="Cu-",]

options(contrasts = c("contr.treatment", "contr.poly"))
model<-glmmTMB(data=df1lm, meanOffFvFm~Apigenin7OGlc*TreatEnv+(1|Plate))
#model<-glmmTMB(data=df1lm, meanOffFvFm~Apigenin7OGlc*TreatEnv+(1|Plate),family=Gamma(log)) #cannot be fixed
sim<-simulateResiduals(model,n=500, plot=T)
summary(model)
options(contrasts = c("contr.sum", "contr.poly"))
modelp<-glmmTMB(data=df1lm, meanOffFvFm~Apigenin7OGlc*TreatEnv+(1|Plate))
car::Anova(modelp,type=3)
plot(allEffects(model))

model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Control",], meanOffFvFm~Apigenin7OGlc)
model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Copper",], meanOffFvFm~Apigenin7OGlc)
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model)

pdf("./FigS6a3.pdf",height=6,width=6)
ggplot(df1lm,aes(y=meanOffFvFm, x=Apigenin7OGlc,col=TreatEnv))+
  geom_point(size=3)+
  scale_color_manual(values = c("Control"="#0056B3", "Copper"="#CC0000"))+
  geom_smooth(method="lm",se=F,size=2)+
  stat_poly_eq(formula = x~y, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  scale_y_continuous(breaks=seq(0.65,0.8,by=0.05),limits = c(0.65,0.8), expand = c(0,0))+
  scale_x_continuous(breaks=seq(0,2,by=0.5),limits = c(0,2), expand = c(0,0))+
  guides(col="none")+
  theme(axis.text.x = element_text(size = 20,color="black", family = "sans"),
        axis.text.y = element_text(size = 20,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```
       # correlation between growth rate under control and plant pre-treatment
```{r}
model<-glmmTMB(data=CtrPreCtr,MApigenin7OGlc~MPreGRArea)
plotResiduals(model,CtrPreCtr$MApigenin7OGlc);plotResiduals(model,CtrPreCtr$MPreGRArea)
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model)

pdf("./FigS4c1.pdf",height=6,width=6)
ggplot(data=CtrPreCtr,aes(x=MPreGRArea,y=MApigenin7OGlc))+
  geom_point(size=3)+
  labs(y="Apigenin7OGlc under control",x="Pre-treatment ratio under control")+
  geom_line(aes(y=predict(model, type="response")),size=1.5,col="#CC0000")+
  scale_x_continuous(breaks=seq(0.9,1.2,by=0.1),limits=c(0.9,1.2),expand=c(0,0))+
  scale_y_continuous(breaks=seq(0.1,0.7,by=0.2),limits=c(0.1,0.7),expand=c(0,0))+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()


model<-glmmTMB(data=CtrPreCu,MApigenin7OGlc~MPreGRArea)
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model)

pdf("./FigS4c2.pdf",height=6,width=6)
ggplot(data=CtrPreCu,aes(x=MPreGRArea,y=MApigenin7OGlc))+
  geom_point(size=3)+
  labs(y="Apigenin7OGlc under control",x="Pre-treatment ratio under copper")+
  geom_line(aes(y=predict(model, type="response")),size=1.5,col="#CC0000")+
  scale_x_continuous(breaks=seq(0.7,1.5,by=0.2),limits=c(0.7,1.5),expand=c(0,0))+
  scale_y_continuous(breaks=seq(0.1,0.7,by=0.2),limits=c(0.1,0.7),expand=c(0,0))+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()
```


  ##Apigenin8CGlc
     #Correlation with mean population growth rate in naive plants
```{r}
df1lm<-df_Met[complete.cases(df_Met$Apigenin8CGlc),]
df1lm<-df1lm[df1lm$Pretreat=="Cu-",]

options(contrasts = c("contr.treatment", "contr.poly"))
model<-glmmTMB(data=df1lm, meanGRArea~Apigenin8CGlc*TreatEnv+(1|Plate))
sim<-simulateResiduals(model,n=500, plot=T)
model<-glmmTMB(data=df1lm, meanGRArea~Apigenin8CGlc*TreatEnv+(1|Plate),contrasts=list(TreatEnv="contr.sum"))
car::Anova(model,type=3)
summary(model)
model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Control",], meanGRArea~Apigenin8CGlc+(1|Plate))
model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Copper",], meanGRArea~Apigenin8CGlc+(1|Plate))
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model,type=2)
plot(allEffects(model))

pdf("./FigS6b4.pdf",height=6,width=6)
ggplot(df1lm,aes(y=meanGRArea, x=Apigenin8CGlc,col=TreatEnv))+
  geom_point(size=3)+
  scale_color_manual(values = c("Control"="#0056B3", "Copper"="#CC0000"))+
  geom_smooth(method="lm",se=F,size=3)+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  scale_y_continuous(breaks=seq(0.2,0.5,by=0.1),limits = c(0.2,0.5), expand = c(0,0))+
  scale_x_continuous(breaks=seq(0,4,by=1),limits = c(0,4), expand = c(0,0))+
  guides(col="none")+
  theme(axis.text.x = element_text(size = 20,color="black", family = "sans"),
        axis.text.y = element_text(size = 20,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```
    #Correlation with meanOffFvFm
```{r}
df1lm<-df_Met[complete.cases(df_Met$Apigenin8CGlc),]
df1lm<-df1lm[complete.cases(df1lm$meanOffFvFm),]
df1lm<-df1lm[df1lm$Pretreat=="Cu-",]

options(contrasts = c("contr.treatment", "contr.poly"))
model<-glmmTMB(data=df1lm, meanOffFvFm~Apigenin8CGlc*TreatEnv+(1|Plate))
model<-glmmTMB(data=df1lm, meanOffFvFm~Apigenin8CGlc*TreatEnv+(1|Plate),family=Gamma(log))
sim<-simulateResiduals(model,n=500, plot=T)
summary(model)
options(contrasts = c("contr.sum", "contr.poly"))
modelp<-glmmTMB(data=df1lm, meanOffFvFm~Apigenin8CGlc*TreatEnv+(1|Plate),family=gaussian(inverse))
car::Anova(modelp,type=3)
plot(allEffects(model))

model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Control",], meanOffFvFm~Apigenin8CGlc)
model<-glmmTMB(data=df1lm[df1lm$TreatEnv=="Copper",], meanOffFvFm~Apigenin8CGlc)
sim<-simulateResiduals(model,n=500, plot=T)
summary(model)
car::Anova(model)

pdf("./FigS6a4.pdf",height=6,width=6)
ggplot(df1lm,aes(y=meanOffFvFm, x=Apigenin8CGlc,col=TreatEnv))+
  geom_point(size=3)+
  scale_color_manual(values = c("Control"="#0056B3", "Copper"="#CC0000"))+
  geom_smooth(method="lm",se=F,size=2)+
  stat_poly_eq(formula = x~y, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  scale_y_continuous(breaks=seq(0.65,0.8,by=0.05),limits = c(0.65,0.8), expand = c(0,0))+
  scale_x_continuous(breaks=seq(0,4,by=1),limits = c(0,4), expand = c(0,0))+
  guides(col="none")+
  theme(axis.text.x = element_text(size = 20,color="black", family = "sans"),
        axis.text.y = element_text(size = 20,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```
       # correlation between growth rate under control and plant pre-treatment
```{r}
model<-glmmTMB(data=CtrPreCtr,MApigenin8CGlc~MPreGRArea)
plotResiduals(model,CtrPreCtr$MApigenin8CGlc);plotResiduals(model,CtrPreCtr$MPreGRArea)
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model)

pdf("./FigS4c3.pdf",height=6,width=6)
ggplot(data=CtrPreCtr,aes(x=MPreGRArea,y=MApigenin8CGlc))+
  geom_point(size=3)+
  labs(y="Apigenin8CGlc under control",x="Pre-treatment ratio under control")+
  geom_line(aes(y=predict(model, type="response")),size=1.5,col="#CC0000")+
  scale_x_continuous(breaks=seq(0.9,1.2,by=0.1),limits=c(0.9,1.2),expand=c(0,0))+
  scale_y_continuous(breaks=seq(0,3,by=1),limits=c(0,3),expand=c(0,0))+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()

model<-glmmTMB(data=CtrPreCu,MApigenin8CGlc~MPreGRArea)
sim<-simulateResiduals(model,n=500, plot=T)
car::Anova(model)

pdf("./FigS4c4.pdf",height=6,width=6)
ggplot(data=CtrPreCu,aes(x=MPreGRArea,y=MApigenin8CGlc))+
  geom_point(size=3)+
  labs(y="Apigenin8CGlc under control",x="Pre-treatment ratio under copper")+
  geom_line(aes(y=predict(model, type="response")),size=1.5,col="#CC0000")+
  scale_x_continuous(breaks=seq(0.7,1.5,by=0.2),limits=c(0.7,1.5),expand=c(0,0))+
  scale_y_continuous(breaks=seq(0,3,by=1),limits=c(0,3),expand=c(0,0))+
  stat_poly_eq(formula = y~x, aes(label = paste(after_stat(rr.label))), parse = TRUE,size=5)+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
      axis.text.y = element_text(size = 18,color="black", family = "sans"),
      axis.title =element_text(size = 25,color="black",face = "bold", family = "sans"),
      axis.ticks = element_line(color="black"))
dev.off()
```


##################################################
##################################################
##################################################
########## Copper concentrations #################
##################################################
##################################################
##################################################

#upload data of metabolites
```{r}
df_Copper<-read_excel("./GeneticBase_TransgnerationalPlasticity.xlsx", col_names = T, na = "",sheet="Copper")
```

# convert to factors
```{r}
df_Copper$Pretreat<-factor(df_Copper$Pretreat)
df_Copper$Pretreat1<-factor(df_Copper$Pretreat, levels=c("Cu-","Cu+"), ordered=TRUE)
df_Copper$Genotype<-factor(df_Copper$Genotype)
df_Copper$Column<-factor(df_Copper$Column)
df_Copper$Row<-factor(df_Copper$Row)
df_Copper$Plate<-factor(df_Copper$Plate)
```

```{r}
df1Cu<-df_Copper[complete.cases(df_Copper$CuConcentration),]

options(contrasts = c("contr.treatment", "contr.poly"))
model1<-glmmTMB(data=df1Cu, CuConcentration~Pretreat+(1|Genotype)+(1|Column))
sim1<-simulateResiduals(model1,n=500, plot=T);testDispersion(sim1)
summary(model1)
car::Anova(model1,type="II")

options(contrasts = c("contr.sum", "contr.poly"))
modelp<-glmmTMB(data=df1Cu, Cyanidine3GlcHPLC~Pretreat*TreatEnv1+(1|Genotype),family=gaussian(inverse),dispformula=~TreatEnv)
car::Anova(modelp,type="III")

dflm1<-(df1Cu %>% dplyr::group_by(Pretreat1) %>% dplyr::summarise(
  CuConc  = mean(CuConcentration),
  se = (sd(CuConcentration)/sqrt(length(CuConcentration)))))

pdf("./FigS8.pdf",height=6,width=4)
ggplot(df1Cu,aes(Pretreat1, CuConcentration))+
  geom_col(data=dflm1,aes(x=Pretreat1, y=CuConc,fill=Pretreat1),width=0.8)+
  geom_errorbar(data=dflm1,aes(x=Pretreat1, y=CuConc,ymin=CuConc-se, ymax=CuConc+se), width=0.3, position=position_dodge(0.80))+
  scale_color_manual(values = c("Cu-"="#0056B3", "Cu+"="#CC0000"))+
  scale_fill_manual(values = c("Cu-"="#0056B3", "Cu+"="#CC0000"))+
  scale_y_continuous(breaks=seq(0,9,by=3),limits = c(0,9), expand = c(0,0))+
  guides(col="none",fill="none")+labs(y="Copper (µg dL-1 mg-1 FW)")+
  theme(axis.text.x = element_text(size = 18,color="black", family = "sans"),
        axis.text.y = element_text(size = 25,color="black", family = "sans"),
        axis.title =element_text(size = 30,color="black",face = "bold", family = "sans"),
        legend.text = element_text(size = 20,color="black", family = "sans"),
        legend.background = element_rect(fill = F),
        legend.key = element_rect(fill = F),
        legend.title = element_text(size = 30,color="black", family = "sans"),
        axis.ticks = element_line(color="black"))
dev.off()
```
